### 2024-09-21 Code/Paper_analysisandtables.Rmd
Create new cleaner project for paper code

---
title: "Paper_analysisandtables"
author: "Kristin Dobbin"
date: "2024-09-18"
output: 
  html_document: default
  pdf_document: default
knit: (function(rmdfile, ...) { rmarkdown::render(rmdfile, output_dir=here::here("Docs"))})
---

```{r setup, include=FALSE, echo=FALSE}
#MARKDOWN SETTINGS
knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Box Sync/Projects/R_Projects/Votingrights_systemperformance", warning=FALSE, message=FALSE, include=FALSE)

#LOAD LIBRARIES
library(tidyverse)
library(vtable)
library(sjPlot)
library(stargazer)

#LOAD AND READY DATA
Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
Data <- Data[,-1]
Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
Data$Final_inst_update <- as.factor(Data$Final_inst_update)

#correct data types, levels and reference factors
Data[,c(1:11,16, 19:44, 46:61)] <- lapply(Data[,c(1:11,16, 19:44, 46:61)], factor)
levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")

Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
Data$FAILING <- as.factor(Data$FAILING)

Data$Source <- as.factor(Data$Source)
Data$public <- as.factor(Data$public)

#re-order factor levels for enfranchisement to be full, limited, none
Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))

Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")

#Make self produced water the reference (compared to purchased)
Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")

Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
Data$CONSTITUENTS_OF_EMERGING_CONCERN_RISK_LEVEL_BI <- relevel(Data$CONSTITUENTS_OF_EMERGING_CONCERN_RISK_LEVEL_BI, ref = "NONE")
Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI <- relevel(Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI, ref = "NONE")
Data$ABESENCE_OF_INTERTIES_RISK_LEVEL <- relevel(Data$ABESENCE_OF_INTERTIES_RISK_LEVEL, ref = "NONE")
Data$BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL <- relevel(Data$BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL, ref = "NONE")
Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL <- relevel(Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL, ref = "NONE")
Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI <- relevel(Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI, ref = "NONE")
Data$EXTREME_WATER_BILL_RISK_LEVEL_BI <- relevel(Data$EXTREME_WATER_BILL_RISK_LEVEL_BI, ref = "NONE")
Data$FUNDING_any_failingoratriskonly <- relevel(Data$FUNDING_any_failingoratriskonly, ref = "No")
Data$OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL <- relevel(Data$OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL <- relevel(Data$MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$DAYS_CASH_ON_HAND_RISK_LEVEL <- relevel(Data$DAYS_CASH_ON_HAND_RISK_LEVEL, ref = "NONE")
Data$OPERATING_RATIO_RISK_LEVEL <- relevel(Data$OPERATING_RATIO_RISK_LEVEL, ref = "NONE")

#add in demographic data
Demographics <- read.csv(here::here("Data_processed/Systemdemographicdata.csv"))
Demographics <- Demographics[,-c(1,3)]
Demographics$PWSID <- as.factor(Demographics$PWSID)

Data <- left_join(Data, Demographics, by = "PWSID")

#MODELS

#Water quality
WQ1 <- glm(WQ_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(WQ1)

##WQ sub components
ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(ecoli)

TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(TT)

MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(MCL)

#Accessibility
Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Accessibility1)

##Accessibility sub components
sources <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(sources)

Interties <- glm(ABESENCE_OF_INTERTIES_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Interties)

Bottled <- glm(BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Bottled)

#Affordability 
Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Affordability1)

## Affordability sub components
MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(MHI)

extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(extreme)

covid <- glm(didnotapplycovid ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(covid)

# TMF
TMF1 <- glm(TMF_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(TMF1)

## TMF sub components
Opcert_violations <- glm(OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Opcert_violations)

mrviolations <- glm(MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(mrviolations)

cash <- glm(DAYS_CASH_ON_HAND_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(cash)

```

This Rmd created html versions of the model results tables using stargazer. The models script must be run prior to using this Rmd

```{r table composites, results='asis', echo=FALSE}
#Composites
WQ1OR <- WQ1
WQ1OR$coefficients <- exp(WQ1OR$coefficients)
p.valuesWQ1 <- list(summary(WQ1)$coefficients[,4])

Affordability1OR <- Affordability1
Affordability1OR$coefficients <- exp(Affordability1OR$coefficients)
pvaluesAffordability1 <- list(summary(Affordability1)$coefficients[,4])

Accessibility1OR <- Accessibility1
Accessibility1OR$coefficients <- exp(Accessibility1OR$coefficients)
pvaluesAccesibility1 <- list(summary(Accessibility1)$coefficients[,4])

TMF1OR <- TMF1
TMF1OR$coefficients <- exp(TMF1OR$coefficients)
pvaluesTMF1 <- list(summary(TMF1)$coefficients[,4])

p.values2 <- list(c(1,1,1,1,1,1))

compositemodellist <- list(WQ1, WQ1OR, Affordability1, Affordability1OR,
                           Accessibility1, Accessibility1OR, TMF1, TMF1OR)

Compositestable <- capture.output(stargazer(compositemodellist,
                                 type = 'html', 
                                 column.labels = c("coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("Water quality", "Affordability", 
                                                    "Accessibility", "TMF"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 p = c(p.valuesWQ1, p.values2, pvaluesAffordability1, p.values2,
                                       pvaluesAccesibility1, p.values2, pvaluesTMF1, p.values2),
                                 omit.stat = c("ll", "aic"),
                                 covariate.labels = c("Lim enfran.", "No enfran.",
                                                      "Population (log)", "Surface water",
                                                      "Purchased"),
                                 omit = "Constant",
                                apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", Compositestable), collapse = "\n"), "\n")
```

<br>
<br>

```{r table WQ, results='asis', echo=FALSE}
#WQ indicators
ecoliOR <- ecoli
ecoliOR$coefficients <- exp(ecoliOR$coefficients)
pvaluesecoli <- list(summary(ecoli)$coefficients[,4])

TTOR <- TT
TTOR$coefficients <- exp(TTOR$coefficients)
pvaluesTT <- list(summary(TT)$coefficients[,4])

MCLOR <- MCL
MCLOR$coefficients <- exp(MCLOR$coefficients)
pvaluesMCL <- list(summary(MCL)$coefficients[,4])

WQmodellist <- list(ecoli, ecoliOR, TT, TTOR, MCL, MCLOR)

WQtable <- capture.output(stargazer(WQmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("ecoli", "treatment technique violations", "MCL violations", 
                                                    "cont. of emerging concern"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 p = c(pvaluesecoli, p.values2, pvaluesTT, p.values2,
                                       pvaluesMCL, p.values2),
                                 omit.stat = c("ll", "aic"),
                                 covariate.labels = c("Lim enfran.", "No enfran.",
                                                      "Population (log)", "Surface water",
                                                      "Purchased"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", WQtable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table affordability, results='asis', echo=FALSE}
#affordability indicators
MHIOR <- MHI
MHIOR$coefficients <- exp(MHIOR$coefficients)
pvaluesMHI <- list(summary(MHI)$coefficients[,4])

extremeOR <- extreme
extremeOR$coefficients <- exp(extremeOR$coefficients)
pvaluesextreme <- list(summary(extreme)$coefficients[,4])

covidOR <- covid
covidOR$coefficients <- exp(covidOR$coefficients)
pvaluescovid <- list(summary(covid)$coefficients[,4])

AFmodellist <- list(MHI, MHIOR, extreme, extremeOR, covid, covidOR)

AFtable <- capture.output(stargazer(AFmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("bill ≥1.5% MHI", "bill ≥150% statewide average",
                                                    "did not apply for arrearage relief", "no funding"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 p = c(pvaluesMHI, p.values2, pvaluesextreme, p.values2, pvaluescovid, p.values2),
                                 omit.stat = c("ll", "aic"),
                                 covariate.labels = c("Lim enfran.", "No enfran.",
                                                      "Population (log)", "Surface water",
                                                      "Purchased"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", AFtable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table accessibility, results='asis', echo=FALSE}
#accessibility indicators
sourcesOR <- sources
sourcesOR$coefficients <- exp(sourcesOR$coefficients)
pvaluessources <- list(summary(sources)$coefficients[,4])

IntertiesOR <- Interties
IntertiesOR$coefficients <- exp(IntertiesOR$coefficients)
pvaluesInterties <- list(summary(Interties)$coefficients[,4])

BottledOR <- Bottled
BottledOR$coefficients <- exp(BottledOR$coefficients)
pvaluesBottled <- list(summary(Bottled)$coefficients[,4])

ACmodellist <- list(sources, sourcesOR, Interties, IntertiesOR, Bottled, BottledOR)

ACtable <- capture.output(stargazer(ACmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("single source", "Absence of interties", "Bottled water reliance"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 p = c(pvaluessources, p.values2, pvaluesInterties, p.values2, pvaluesBottled, p.values2),
                                 omit.stat = c("ll", "aic"),
                                 covariate.labels = c("Lim enfran.", "No enfran.",
                                                      "Population (log)", "Surface water",
                                                      "Purchased"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", ACtable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table TMF, results='asis', echo=FALSE}
#TMF indicators
Opcert_violationsOR <- Opcert_violations
Opcert_violationsOR$coefficients <- exp(Opcert_violationsOR$coefficients)
pvaluesOpcert <- list(summary(Opcert_violations)$coefficients[,4])

mrviolationsOR <- mrviolations
mrviolationsOR$coefficients <- exp(mrviolationsOR$coefficients)
pvaluesmrviolations <- list(summary(mrviolations)$coefficients[,4])

cashOR <- cash
cashOR$coefficients <- exp(cashOR$coefficients)
pvaluescash <- list(summary(cash)$coefficients[,4])

TMFmodellist <- list(Opcert_violations, Opcert_violationsOR, mrviolations, mrviolationsOR, cash, cashOR)

TMFtable <- capture.output(stargazer(TMFmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio", 
                                                   "coefficients(se)", "Odds ratio"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("operator cert violations", "M&R violations", "Cash on hand"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 p = c(pvaluesOpcert, p.values2, pvaluesmrviolations, p.values2, pvaluescash, p.values2),
                                 omit.stat = c("ll", "aic"),
                                 covariate.labels = c("Lim enfran.", "No enfran.",
                                                      "Population (log)", "Surface water",
                                                      "Purchased"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", TMFtable), collapse = "\n"), "\n")
```

<br>
<br>

### 2024-09-21 Code/Data_compilation_systemperformance.R

@@ -1,0 +1,349 @@
+#Compile base data
+
+#load libraries
+library(tidyverse)
+library(plyr)
+
+# Create sample and governance variables ----------------------------------------------------
+
+###Load data and initial manipulations
+CWSgov <- read.csv(here::here("Data_raw/Dobbin_Fencl_McBride_2023_CA_CWS_institutional_type_update.csv"), header=T,
+                   na.strings=c("","NA"))
+CWSgov$Primacy_FINAL <- as.factor(CWSgov$Primacy_FINAL)
+CWSgov <- CWSgov %>% filter(Primacy_FINAL != "EPA Region 9") #get rid of EPA regulated Tribal water systems
+CWSgov <- CWSgov %>% filter(Final_inst_update != "Tribal Government") #Get rid of three remaining systems that are Tribal but state regulated
+CWSgov$Primacy_FINAL <- droplevels(CWSgov$Primacy_FINAL)
+CWSgov$PWSID <- as.factor(CWSgov$PWSID)
+CWSgov$Final_inst_update <- as.factor(CWSgov$Final_inst_update)
+CWSgov$Inst_Subtype  <- as.factor(CWSgov$Inst_Subtype)
+
+### create new variables for ancillary/non ancillary, public/private and customer enfranchisement
+
+## To focus on residential systems where residents are served and could have a role in governance/decision-making, identify which systems are "residential" this is mostly synonymous to non-ancillary except with this variable maintaining Mobile Home Parks which are generally considered to be ancillary and yet in other ways these systems are more similar to non-ancillary systems in that they serve a stable residential base. Moreover, California has seen significant organizing among Mobile Home Park communities in recent years around drinking water issues. 
+CWSgov$Residential <- ifelse(CWSgov$Final_inst_update == "Community Services District" |
+                               CWSgov$Final_inst_update == "County Service Area" | 
+                               CWSgov$Final_inst_update == "Municipal Water District" | 
+                               CWSgov$Final_inst_update == "City" | 
+                               CWSgov$Final_inst_update == "County Sanitation District"| 
+                               CWSgov$Final_inst_update == "County Water District"| 
+                               CWSgov$Final_inst_update == "County Waterworks District" |
+                               CWSgov$Final_inst_update == "Irrigation District"| 
+                               CWSgov$Final_inst_update == "Maintenance District"|
+                               CWSgov$Final_inst_update == "Municipal Utility District"|
[…]

### 2024-09-21 Code/Data_compiliation_SES.R

@@ -1,0 +1,82 @@
+#Data compilation for demographics
+
+#LOAD LIBRARIES
+library(tidyverse)
+library(ggmap)
+library(censusapi)
+library(tidycensus)
+library(sf)
+library(sp)
+
+Demographics_blockg <- get_acs(geography = "block group", variables = c("B03002_001E", "B03002_003E", "B03002_004E", "B03002_005E", "B03002_006E", "B03002_007E", "B03002_012E", "B19013_001E", "B25003_001E", "B25003_002E", "B25003_003E", "B11001_001E"), state = "CA", year =2020, output = "wide", survey = "acs5", geometry = TRUE)
+
+#rename variables
+Demographics_blockg <- Demographics_blockg %>% dplyr::rename(Race.estimate.total = B03002_001E, Race.white.alone = B03002_003E, Race.black.alone = B03002_004E, Race.native.alone = B03002_005E, Race.asian.alone = B03002_006E, Race.PI.alone = B03002_007E, Race.hispanicorlatino = B03002_012E, Median.hh.income = B19013_001E, Tenure.estimate.total = B25003_001E, Tenure.owner = B25003_002E, Tenure.renter = B25003_003E, Households.total = B11001_001E)
+
+#Make median household income into a count by multiplying median income by total households (see https://crd230.github.io/lab3.html)
+Demographics_blockg$Income.aggregatecount <- Demographics_blockg$Median.hh.income*Demographics_blockg$Households.total
+
+#load in boundary polygons for CWS
+PWS_boundary <- st_read("Data_raw/PWS/") #Consider updating these soon, this file is a year or two old at this point
+str(PWS_boundary)
+#plot(st_geometry(PWS_boundary)) #takes a while but works!
+
+PWS_boundary$SABL_PWSID <- as.factor(PWS_boundary$SABL_PWSID)
+PWS_boundary$WATER_SYST <- as.factor(PWS_boundary$WATER_SYST)
+PWS_boundary <- PWS_boundary %>% filter(BOUNDARY_T != "Jurisdictional") #remove jurisdictional boundaries to get rid of duplicates
+
+Data_small <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
+Data_small <- Data_small[,c(2,11)]
+  
+Data_geo <- left_join(PWS_boundary, Data_small, by = c("SABL_PWSID" = "PWSID"))
[…]

### 2024-09-21 Code/Figures.PP.R

@@ -1,0 +1,155 @@
+#Predicted probabilities figures for GLMs. NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Models.R script
+library(marginaleffects)
+
+#pull out probabilities in confidence intervals
+Effect_WQ <- predictions(WQ1,
+                         newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+Effect_Accessibility <- predictions(Accessibility1,
+                                    newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+Effect_Affordability <- predictions(Affordability1,
+                                    newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+Effect_TMF <- predictions(TMF1,
+                          newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+
+#add key and combine data sets
+Effect_WQ$Key <- "Water quality risk"
+Effect_Accessibility$Key <- "Accessibility risk"
+Effect_Affordability$Key <- "Affordability risk"
+Effect_TMF$Key <- "TMF capacity risk"
+Probabilities <- bind_rows(Effect_Accessibility, Effect_Affordability, Effect_TMF, Effect_WQ)
+
+#make figure faceting with key
+plotCombined <- Probabilities %>%
+  ggplot(aes(x = enfranchisement_final, y = estimate)) +
+  geom_errorbar(aes(ymin = conf.low, ymax = conf.high, color = enfranchisement_final), width = 0.2, show.legend = FALSE) +
+  geom_point(aes(color = enfranchisement_final), show.legend = FALSE) +
+  xlab(" ") +
+  ylab(" ") +
+  scale_color_brewer(palette = "Dark2") +
+  facet_wrap(~Key, scales = "free", nrow = 2, ncol = 2) +
+  cowplot::theme_half_open(); plotCombined
+
+ggsave("Fig2_tmp.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
+
+#water quality subindicators figure
+Effect_WQ_ecoli <- predictions(ecoli,
+                               newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+Effect_WQ_TT <- predictions(TT,
[…]

### 2024-09-21 Code/Figures_Map.R

@@ -1,0 +1,32 @@
+#Map systems by enfranchisement type. NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Data_compillation_SES.R script
+
+#Load libraries
+library(tidyverse)
+library(sf)
+library(ggmap)
+library(mapproj)
+library(maps)
+
+#load data 
+Data_geo$Centroid <- NA
+Data_geo$Centroid <- st_centroid(Data_geo$geometry) #Get centroids to service area boundaries
+Data_geo <- Data_geo %>% drop_na(enfranchisement_final) #get rid of NAs. Mostly cuts out all the non-CWS but also some CWS in the analysis without service area boundary data
+
+#Basemap
+caCountiesTmp <- tigris::counties(state = 06) %>%
+  st_as_sf()
+
+#maps
+Enfranchisement_Type <- ggplot() +
+  geom_sf(data = caCountiesTmp) +
+  geom_sf(mapping = aes(colour = enfranchisement_final, geometry = Centroid), data = Data_geo, inherit.aes = FALSE) +
+  theme_void() +
+  theme(legend.position = "bottom", 
+        legend.margin=margin(0,0,0,0),
+        legend.box.margin=margin(10,10,10,10)) +
+  labs(color = "Enfranchisement type") +
+  theme(panel.background=element_blank())
+
+ggsave('Figures/Map.png', width = 4, height = 6, dpi = 300, bg='#ffffff')
+
+Tulare <- Data_geo %>% filter(COUNTY == "TULARE")

### 2024-09-21 Code/Models.R

@@ -1,0 +1,113 @@
+# Models (This code should match Paper_analysisandtables.Rmd)
+
+#LOAD LIBRARIES
+library(tidyverse)
+library(vtable)
+library(sjPlot)
+
+#LOAD AND READY DATA
+Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
+Data <- Data[,-1]
+Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
+Data$Final_inst_update <- as.factor(Data$Final_inst_update)
+
+#correct data types, levels and reference factors
+Data[,c(1:11,16, 19:44, 46:61)] <- lapply(Data[,c(1:11,16, 19:44, 46:61)], factor)
+levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")
+
+Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
+Data$FAILING <- as.factor(Data$FAILING)
+
+Data$Source <- as.factor(Data$Source)
+Data$public <- as.factor(Data$public)
+
+#re-order factor levels for enfranchisement to be full, limited, none
+Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))
+
+Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
+Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")
+
+#Make self produced water the reference (compared to purchased)
+Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")
+
+Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
+Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
+Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
[…]

### 2024-11-15 Code/Paper_analysisandtables.Rmd
Catch up

@@
 #MARKDOWN SETTINGS
-knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Box Sync/Projects/R_Projects/Votingrights_systemperformance", warning=FALSE, message=FALSE, include=FALSE)
+#knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Documents/Projects/R_Projects/Votingrights_systemperformance", warning=FALSE, message=FALSE, include=FALSE)
 
@@
 compositemodellist <- list(WQ1, WQ1OR, Affordability1, Affordability1OR,
-                           Accessibility1, Accessibility1OR, TMF1, TMF1OR)
+                          TMF1, TMF1OR)
 
@@
                                                    "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
                                                    "coefficients(se)", "Odds ratio"),
@@
                                  dep.var.labels = c("Water quality", "Affordability", 
-                                                    "Accessibility", "TMF"), 
-                                 ci=c(F,T, F, T, F, T, F, T), 
+                                                    "TMF"), 
+                                 ci=c(F,T, F, T, F, T), 
                                  star.cutoffs = c(0.05, 0.01, 0.001),
                                  p = c(p.valuesWQ1, p.values2, pvaluesAffordability1, p.values2,
-                                       pvaluesAccesibility1, p.values2, pvaluesTMF1, p.values2),
+                                       pvaluesTMF1, p.values2),
                                  omit.stat = c("ll", "aic"),
@@
 <br>
+
+```{r descriptives for demofraphics, results='asis', echo=FALSE}
+Data_mini <- Data %>% select(enfranchisement_final, MHI, MHI_KD, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, TMF_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian)
+
+Descriptives<- Data_mini %>%
+  group_by(enfranchisement_final) %>% 
+  summarize(meanLNPOP = mean(na.omit(POPULATION)),
+            meanMHI = mean(na.omit(MHI)),
+            meanPercRent = mean(na.omit(Percent.renter)),
+            meanPercWhite = mean(na.omit(Percent.white)),
+            meanCES = mean(na.omit(CALENVIRO_SCREEN_SCORE)),
+            medianpop = median(na.omit(POPULATION)))
+
+SIdescriptives <- Data_mini %>% 
+  group_by(enfranchisement_final) %>% 
+  select_if(is.numeric) %>% 
+  gather(Variable, Value, -enfranchisement_final) %>% 
+  group_by(Variable, enfranchisement_final) %>% 
+  summarise(min = min(na.omit(Value)), 
+            median = median(na.omit(Value)),
+            mean = mean(na.omit(Value)), 
+            max = max(na.omit(Value)),
+            sd = sd(na.omit(Value))) %>% 
+ ungroup()
+
+SIdescriptives <- SIdescriptives %>% mutate_if(is.numeric, round, digits=3)
+```

### 2024-11-15 Code/Data_compilation_systemperformance.R

@@ -7,8 +7,7 @@ library(plyr)
 # Create sample and governance variables ----------------------------------------------------
 
 ###Load data and initial manipulations
-CWSgov <- read.csv(here::here("Data_raw/Dobbin_Fencl_McBride_2023_CA_CWS_institutional_type_update.csv"), header=T,
-                   na.strings=c("","NA"))
+CWSgov <- read.csv(here::here("Data_raw/Dobbin_Fencl_McBride_2023_CA_CWS_institutional_type_update.csv"), header=T, na.strings=c("","NA"))
 CWSgov$Primacy_FINAL <- as.factor(CWSgov$Primacy_FINAL)
 CWSgov <- CWSgov %>% filter(Primacy_FINAL != "EPA Region 9") #get rid of EPA regulated Tribal water systems
 CWSgov <- CWSgov %>% filter(Final_inst_update != "Tribal Government") #Get rid of three remaining systems that are Tribal but state regulated

### 2024-11-15 Code/Figures_Map.R

@@ -6,6 +6,7 @@ library(sf)
 library(ggmap)
 library(mapproj)
 library(maps)
+library(gridExtra)
 
 #load data 
 Data_geo$Centroid <- NA
@@ -25,8 +26,66 @@ Enfranchisement_Type <- ggplot() +
         legend.margin=margin(0,0,0,0),
         legend.box.margin=margin(10,10,10,10)) +
   labs(color = "Enfranchisement type") +
+  scale_color_brewer(palette = "Dark2") +
   theme(panel.background=element_blank())
 
 ggsave('Figures/Map.png', width = 4, height = 6, dpi = 300, bg='#ffffff')
 
+#make smaller zoomed in demo map 1
+caCountiesTmp_Tulare <- caCountiesTmp %>% filter(NAME == "Tulare")
+
 Tulare <- Data_geo %>% filter(COUNTY == "TULARE")
+ExeterandTooleville <- Tulare %>% filter(SABL_PWSID == "CA5410003" | SABL_PWSID == "CA5400567")
+ExeterandTooleville$Label <- c("City of Exeter", "Tooleville Mutual Water Company")
+
+Demo1 <- ggplot() +
+  geom_sf(mapping = aes(colour = enfranchisement_final, geometry = geometry), data = ExeterandTooleville, 
+          inherit.aes = FALSE) +
+  theme_void() +
+  theme(legend.position = "bottom", 
+        legend.margin=margin(0,0,0,0),
+        legend.box.margin=margin(10,10,10,10)) +
+  labs(color = "Enfranchisement type") +
+  scale_color_brewer(palette = "Dark2") +
+  geom_sf_text(data = ExeterandTooleville, aes(label = Label))
+  theme(panel.background=element_blank()); Demo1
+
+ggsave('Figures/Demo_map1.png', width = 6, height = 3, dpi = 300, bg='#ffffff')
+
+#make smaller zoomed in demo map 2
+Kern <- Data_geo %>% filter(COUNTY == "KERN")
+Kerndemo <- Kern %>% filter(SABL_PWSID == "CA1500588" | SABL_PWSID == "CA1510001")
+Kerndemo$Label <- c("Arvin Community Services District", "Son Shine Properties")
+
+Demo2 <- ggplot() +
+  geom_sf(mapping = aes(colour = enfranchisement_final, geometry = geometry), data = Kerndemo, 
+          inherit.aes = FALSE) +
+  theme_void() +
[…]

### 2024-11-15 Code/Figures_PP.R

@@ -16,17 +16,17 @@ Effect_WQ$Key <- "Water quality risk"
 Effect_Accessibility$Key <- "Accessibility risk"
 Effect_Affordability$Key <- "Affordability risk"
 Effect_TMF$Key <- "TMF capacity risk"
-Probabilities <- bind_rows(Effect_Accessibility, Effect_Affordability, Effect_TMF, Effect_WQ)
+Probabilities <- bind_rows(Effect_WQ, Effect_Affordability, Effect_TMF)
 
 #make figure faceting with key
 plotCombined <- Probabilities %>%
-  ggplot(aes(x = enfranchisement_final, y = estimate)) +
-  geom_errorbar(aes(ymin = conf.low, ymax = conf.high, color = enfranchisement_final), width = 0.2, show.legend = FALSE) +
+  ggplot(aes(x = enfranchisement_final, y = estimate*100)) +
+  geom_errorbar(aes(ymin = conf.low*100, ymax = conf.high*100, color = enfranchisement_final), width = 0.2, show.legend = FALSE) +
   geom_point(aes(color = enfranchisement_final), show.legend = FALSE) +
   xlab(" ") +
   ylab(" ") +
   scale_color_brewer(palette = "Dark2") +
-  facet_wrap(~Key, scales = "free", nrow = 2, ncol = 2) +
+  facet_wrap(~Key, scales = "free", nrow = 1, ncol = 3) +
   cowplot::theme_half_open(); plotCombined
 
 ggsave("Fig2_tmp.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
@@ -131,13 +131,13 @@ plotCombined_TMF <- Probabilities_TMF %>%
 library(RColorBrewer)
 
 #Probabilities_MASSIVE <- bind_rows(Effect_WQ_ecoli, Effect_WQ_CEC, Effect_WQ_MCL, Effect_WQ_TT, Effect_AF_MHI, Effect_AF_extreme, Effect_AF_covid, Effect_AF_funding, Effect_AC_sources, Effect_AC_interties, Effect_AC_bottled, Effect_AC_sourcecapacity, Effect_TMF_opcertviolations, Effect_TMF_mrviolations, Effect_TMF_cash, Effect_TMF_operating)
-Probabilities_MASSIVE <- bind_rows(Probabilities_WQ, Probabilities_AF, Probabilities_AC, Probabilities_TMF)
[…]

### 2024-11-15 Sensitivitytest_MHP.R

@@ -1,4 +1,4 @@
-# Models (This code should match Paper_analysisandtables.Rmd)
+# Sensitivity analysis making MHPs their own category
 
 #LOAD LIBRARIES
 library(tidyverse)
@@ -52,6 +52,14 @@ Demographics <- Demographics[,-c(1,3)]
 Demographics$PWSID <- as.factor(Demographics$PWSID)
 
 Data <- left_join(Data, Demographics, by = "PWSID")
+
+#Finally make MHP their own category
+#Make Mobile Home Parks (MHP) their own enfranchisement category
+levels(Data$enfranchisement_final) <- c(levels(Data$enfranchisement_final),"Mobile Home Park")
+
+Data <- Data %>% mutate(enfranchisement_final = case_when(Final_inst_update == "Mobile Home Park" ~ 
+                                                            "Mobile Home Park", .default = enfranchisement_final)) %>%
+  mutate(enfranchisement_final = as.factor(enfranchisement_final))
 
 #MODELS
 #Water quality

### 2024-12-10 Code/Paper_analysisandtables.Rmd
Updating to 2022 ACS data and updated PWS boundary data, revise Fig 1 (map)

@@
 
-#Accessibility
-Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(Accessibility1)
-
-##Accessibility sub components
-sources <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(sources)
-
-Interties <- glm(ABESENCE_OF_INTERTIES_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(Interties)
-
-Bottled <- glm(BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(Bottled)
-
 #Affordability 
@@
 
-Accessibility1OR <- Accessibility1
-Accessibility1OR$coefficients <- exp(Accessibility1OR$coefficients)
-pvaluesAccesibility1 <- list(summary(Accessibility1)$coefficients[,4])
-
 TMF1OR <- TMF1
@@
 cat(paste(gsub("\\(0.000, 0.000\\)", "", Compositestable), collapse = "\n"), "\n")
+
 ```
@@
 
-```{r table accessibility, results='asis', echo=FALSE}
-#accessibility indicators
-sourcesOR <- sources
-sourcesOR$coefficients <- exp(sourcesOR$coefficients)
-pvaluessources <- list(summary(sources)$coefficients[,4])
-
-IntertiesOR <- Interties
-IntertiesOR$coefficients <- exp(IntertiesOR$coefficients)
-pvaluesInterties <- list(summary(Interties)$coefficients[,4])
-
-BottledOR <- Bottled
-BottledOR$coefficients <- exp(BottledOR$coefficients)
-pvaluesBottled <- list(summary(Bottled)$coefficients[,4])
-
-ACmodellist <- list(sources, sourcesOR, Interties, IntertiesOR, Bottled, BottledOR)
-
-ACtable <- capture.output(stargazer(ACmodellist,
-                                 type = 'html', 
-                                 column.labels = c("coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio"),
-                                 model.numbers = FALSE, 
-                                 dep.var.labels = c("single source", "Absence of interties", "Bottled water reliance"), 
-                                 ci=c(F,T, F, T, F, T, F, T), 
-                                 star.cutoffs = c(0.05, 0.01, 0.001),
-                                 p = c(pvaluessources, p.values2, pvaluesInterties, p.values2, pvaluesBottled, p.values2),
-                                 omit.stat = c("ll", "aic"),
-                                 covariate.labels = c("Lim enfran.", "No enfran.",
-                                                      "Population (log)", "Surface water",
-                                                      "Purchased"),
-                                 omit = "Constant",
-                                 apply.ci = function(x) { 0 }))
-
-cat(paste(gsub("\\(0.000, 0.000\\)", "", ACtable), collapse = "\n"), "\n")
-
-```
-
-<br>
-<br>
-
 ```{r table TMF, results='asis', echo=FALSE}
@@
 ```{r descriptives for demofraphics, results='asis', echo=FALSE}
-Data_mini <- Data %>% select(enfranchisement_final, MHI, MHI_KD, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, TMF_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian)
+Data_mini <- Data %>% select(enfranchisement_final, MHI_KD, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, TMF_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian)
 
@@
   group_by(enfranchisement_final) %>% 
-  summarize(meanLNPOP = mean(na.omit(POPULATION)),
+  summarize(meanPOP = mean(na.omit(POPULATION)),
+            meanMHI_KD = mean(na.omit(MHI_KD)),
             meanMHI = mean(na.omit(MHI)),
@@
 ```
+
+```{r dem sig tests, results='asis', echo=FALSE}
+#ANOVA and Tukey HSD
+MHI <- aov(MHI_KD ~ enfranchisement_final, data = Data)
+summary(MHI)
+TukeyMHI <- TukeyHSD(MHI)
+TukeyMHI_table <- as.data.frame(TukeyMHI$enfranchisement_final)
+
+AA <- aov(Percent.black ~ enfranchisement_final, data = Data)
+summary(AA)
+TukeyAA <- TukeyHSD(AA)
+TukeyAA_table <- as.data.frame(TukeyAA$enfranchisement_final)
+```

### 2024-12-10 Code/Data_compiliation_SES.R

@@ -8,7 +8,7 @@ library(tidycensus)
 library(sf)
 library(sp)
 
-Demographics_blockg <- get_acs(geography = "block group", variables = c("B03002_001E", "B03002_003E", "B03002_004E", "B03002_005E", "B03002_006E", "B03002_007E", "B03002_012E", "B19013_001E", "B25003_001E", "B25003_002E", "B25003_003E", "B11001_001E"), state = "CA", year =2020, output = "wide", survey = "acs5", geometry = TRUE)
+Demographics_blockg <- get_acs(geography = "block group", variables = c("B03002_001E", "B03002_003E", "B03002_004E", "B03002_005E", "B03002_006E", "B03002_007E", "B03002_012E", "B19013_001E", "B25003_001E", "B25003_002E", "B25003_003E", "B11001_001E"), state = "CA", year =2022, output = "wide", survey = "acs5", geometry = TRUE)
 
 #rename variables
 Demographics_blockg <- Demographics_blockg %>% dplyr::rename(Race.estimate.total = B03002_001E, Race.white.alone = B03002_003E, Race.black.alone = B03002_004E, Race.native.alone = B03002_005E, Race.asian.alone = B03002_006E, Race.PI.alone = B03002_007E, Race.hispanicorlatino = B03002_012E, Median.hh.income = B19013_001E, Tenure.estimate.total = B25003_001E, Tenure.owner = B25003_002E, Tenure.renter = B25003_003E, Households.total = B11001_001E)
@@ -17,7 +17,7 @@ Demographics_blockg <- Demographics_blockg %>% dplyr::rename(Race.estimate.total
 Demographics_blockg$Income.aggregatecount <- Demographics_blockg$Median.hh.income*Demographics_blockg$Households.total
 
 #load in boundary polygons for CWS
-PWS_boundary <- st_read("Data_raw/PWS/") #Consider updating these soon, this file is a year or two old at this point
+PWS_boundary <- st_read("Data_raw/California_Drinking_Water_System_Area_Boundaries/") #New version of SABL downloaded November 15, 2024
 str(PWS_boundary)
 #plot(st_geometry(PWS_boundary)) #takes a while but works!
 
@@ -29,8 +29,9 @@ Data_small <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
[…]

### 2024-12-10 Code/Data_compiliation_SES_updated2022.R

@@ -1,0 +1,89 @@
+#Data compilation for demographics updated to 2022 five year estimates
+
+#LOAD LIBRARIES
+library(tidyverse)
+library(ggmap)
+library(censusapi)
+library(tidycensus)
+library(sf)
+library(sp)
+
+Demographics_blockg <- get_acs(geography = "block group", variables = c("B03002_001E", "B03002_003E", "B03002_004E", "B03002_005E", "B03002_006E", "B03002_007E", "B03002_012E", "B19013_001E", "B25003_001E", "B25003_002E", "B25003_003E", "B11001_001E"), state = "CA", year =2022, output = "wide", survey = "acs5", geometry = TRUE)
+
+#rename variables
+Demographics_blockg <- Demographics_blockg %>% dplyr::rename(Race.estimate.total = B03002_001E, Race.white.alone = B03002_003E, Race.black.alone = B03002_004E, Race.native.alone = B03002_005E, Race.asian.alone = B03002_006E, Race.PI.alone = B03002_007E, Race.hispanicorlatino = B03002_012E, Median.hh.income = B19013_001E, Tenure.estimate.total = B25003_001E, Tenure.owner = B25003_002E, Tenure.renter = B25003_003E, Households.total = B11001_001E)
+
+#Make median household income into a count by multiplying median income by total households (see https://crd230.github.io/lab3.html)
+Demographics_blockg$Income.aggregatecount <- Demographics_blockg$Median.hh.income*Demographics_blockg$Households.total
+
+#load in boundary polygons for CWS
+PWS_boundary <- st_read("Data_raw/California_Drinking_Water_System_Area_Boundaries/") #New version of SABL downloaded November 15, 2024
+str(PWS_boundary)
+#plot(st_geometry(PWS_boundary)) #takes a while but works!
+
+PWS_boundary$SABL_PWSID <- as.factor(PWS_boundary$SABL_PWSID)
+PWS_boundary$WATER_SYST <- as.factor(PWS_boundary$WATER_SYST)
+PWS_boundary <- PWS_boundary %>% filter(BOUNDARY_T != "Jurisdictional") #remove jurisdictional boundaries to get rid of duplicates
+
+Data_small <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
+Data_small <- Data_small[,c(2,11)]
+
[…]

### 2024-12-10 Code/Figures_Map.R

@@ -19,23 +19,32 @@ caCountiesTmp <- tigris::counties(state = 06) %>%
 
 #maps
 Enfranchisement_Type <- ggplot() +
-  geom_sf(data = caCountiesTmp) +
-  geom_sf(mapping = aes(colour = enfranchisement_final, geometry = Centroid), data = Data_geo, inherit.aes = FALSE) +
+  geom_sf(data = caCountiesTmp, fill = "white") +
+  geom_sf(data = caCountiesTmp_Mendocino, linewidth = .75, fill = "white") +
+  geom_sf(mapping = aes(colour = enfranchisement_final, geometry = Centroid), size = .35, data = Data_geo, inherit.aes = FALSE) +
   theme_void() +
   theme(legend.position = "bottom", 
+        legend.direction = "vertical",
         legend.margin=margin(0,0,0,0),
         legend.box.margin=margin(10,10,10,10)) +
   labs(color = "Enfranchisement type") +
   scale_color_brewer(palette = "Dark2") +
-  theme(panel.background=element_blank())
+  theme(panel.background=element_blank()); Enfranchisement_Type
 
-ggsave('Figures/Map.png', width = 4, height = 6, dpi = 300, bg='#ffffff')
+ggsave('Figures/Map.png', width = 5, height = 8, dpi = 300, bg='#ffffff')
+
+#export county data to fix county variable
+Countydata <- Data_geo %>% select(SABL_PWSID, COUNTY) %>% st_drop_geometry()
+write.csv(Countydata, file = here::here("Data_processed/Countydata.csv"))
+Counties <- read.csv("Data_processed/Countydata_fixed.csv")
+Data_geo <- left_join(Data_geo, Counties, by = "SABL_PWSID")
+Data_geo$COUNTY_FIXED <- as.factor(Data_geo$COUNTY_FIXED)
 
 #make smaller zoomed in demo map 1
 caCountiesTmp_Tulare <- caCountiesTmp %>% filter(NAME == "Tulare")
 
-Tulare <- Data_geo %>% filter(COUNTY == "TULARE")
-ExeterandTooleville <- Tulare %>% filter(SABL_PWSID == "CA5410003" | SABL_PWSID == "CA5400567")
+Tulare <- Data_geo %>% filter(Data_geo$COUNTY_FIXED == "TULARE")
[…]

### 2024-12-10 Code/Figures_PP.R

@@ -4,8 +4,6 @@ library(marginaleffects)
 #pull out probabilities in confidence intervals
 Effect_WQ <- predictions(WQ1,
                          newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
-Effect_Accessibility <- predictions(Accessibility1,
-                                    newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
 Effect_Affordability <- predictions(Affordability1,
                                     newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
 Effect_TMF <- predictions(TMF1,
@@ -13,10 +11,11 @@ Effect_TMF <- predictions(TMF1,
 
 #add key and combine data sets
 Effect_WQ$Key <- "Water quality risk"
-Effect_Accessibility$Key <- "Accessibility risk"
 Effect_Affordability$Key <- "Affordability risk"
 Effect_TMF$Key <- "TMF capacity risk"
-Probabilities <- bind_rows(Effect_WQ, Effect_Affordability, Effect_TMF)
+Probabilities <- bind_rows(Effect_WQ, Effect_TMF, Effect_Affordability)
+Probabilities$Key <- factor(Probabilities$Key, levels = c("Water quality risk", "TMF capacity risk", "Affordability risk"))
+
 
 #make figure faceting with key
 plotCombined <- Probabilities %>%
@@ -29,7 +28,7 @@ plotCombined <- Probabilities %>%
   facet_wrap(~Key, scales = "free", nrow = 1, ncol = 3) +
   cowplot::theme_half_open(); plotCombined
 
-ggsave("Fig2_tmp.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
+ggsave("Fig2.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
 
 #water quality subindicators figure
 Effect_WQ_ecoli <- predictions(ecoli,
@@ -80,30 +79,6 @@ plotCombined_AF <- Probabilities_AF %>%
   facet_wrap(~Key, scale = "free", nrow = 2, ncol = 2) +
   cowplot::theme_half_open(); plotCombined_AF
 
-#accessibility sub indicators figure
-Effect_AC_sources <- predictions(sources,
-                                 newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
[…]

### 2024-12-10 Code/Models.R

@@ -54,7 +54,6 @@ Demographics$PWSID <- as.factor(Demographics$PWSID)
 Data <- left_join(Data, Demographics, by = "PWSID")
 
 #MODELS
-#Water quality
 Failing <- glm(FAILING ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Failing)
 
@@ -71,20 +70,6 @@ summary(TT)
 
 MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MCL)
-
-#Accessibility
-Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(Accessibility1)
-
-##Accessibility sub components
-sources <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(sources)
-
-Interties <- glm(ABESENCE_OF_INTERTIES_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(Interties)
-
-Bottled <- glm(BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(Bottled)
 
 #Affordability 
 Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)

### 2024-12-10 Code/Sensitivity_analysis.R

@@ -1,0 +1,243 @@
+#Sensitivity analysis - breaking out enfranchisement types more
+
+
+#LOAD LIBRARIES
+library(tidyverse)
+library(vtable)
+library(sjPlot)
+
+#LOAD AND READY DATA
+Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
+Data <- Data[,-1]
+Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
+Data$Final_inst_update <- as.factor(Data$Final_inst_update)
+
+#correct data types, levels and reference factors
+Data[,c(1:11,16, 19:44, 46:61)] <- lapply(Data[,c(1:11,16, 19:44, 46:61)], factor)
+levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")
+
+Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
+Data$FAILING <- as.factor(Data$FAILING)
+
+Data$Source <- as.factor(Data$Source)
+Data$public <- as.factor(Data$public)
+
+#re-order factor levels for enfranchisement to be full, limited, none
+Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))
+
+Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
+Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")
+
+#Make self produced water the reference (compared to purchased)
+Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")
+
+Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
+Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
+Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
[…]

### 2024-12-20 Code/Paper_analysisandtables.Rmd
Final figure updates

@@
                                  type = 'html', 
-                                 column.labels = c("coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio"),
+                                 column.labels = c("coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR"),
                                  model.numbers = FALSE, 
                                  dep.var.labels = c("Water quality", "Affordability", 
-                                                    "TMF"), 
+                                                    "TMF capacity"), 
                                  ci=c(F,T, F, T, F, T), 
                                  star.cutoffs = c(0.05, 0.01, 0.001),
+                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
+                                 notes.append = FALSE,
                                  p = c(p.valuesWQ1, p.values2, pvaluesAffordability1, p.values2,
@@
                                  omit.stat = c("ll", "aic"),
-                                 covariate.labels = c("Lim enfran.", "No enfran.",
+                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                       "Population (log)", "Surface water",
-                                                      "Purchased"),
+                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                  omit = "Constant",
@@
                                  type = 'html', 
-                                 column.labels = c("coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio"),
+                                 column.labels = c("coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR"),
                                  model.numbers = FALSE, 
-                                 dep.var.labels = c("ecoli", "treatment technique violations", "MCL violations", 
-                                                    "cont. of emerging concern"), 
+                                 dep.var.labels = c("E. coli", "Treatment technique violations", "MCL violations"), 
                                  ci=c(F,T, F, T, F, T, F, T), 
                                  star.cutoffs = c(0.05, 0.01, 0.001),
+                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
+                                 notes.append = FALSE,
                                  p = c(pvaluesecoli, p.values2, pvaluesTT, p.values2,
@@
                                  type = 'html', 
-                                 column.labels = c("coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio"),
+                                 column.labels = c("coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR"),
                                  model.numbers = FALSE, 
-                                 dep.var.labels = c("bill ≥1.5% MHI", "bill ≥150% statewide average",
-                                                    "did not apply for arrearage relief", "no funding"), 
+                                 dep.var.labels = c("Unaffordable bill", "Extreme bill",
+                                                    "No arrearage relief"), 
                                  ci=c(F,T, F, T, F, T, F, T), 
                                  star.cutoffs = c(0.05, 0.01, 0.001),
+                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
+                                 notes.append = FALSE,
                                  p = c(pvaluesMHI, p.values2, pvaluesextreme, p.values2, pvaluescovid, p.values2),
                                  omit.stat = c("ll", "aic"),
-                                 covariate.labels = c("Lim enfran.", "No enfran.",
+                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                       "Population (log)", "Surface water",
-                                                      "Purchased"),
+                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                  omit = "Constant",
@@
                                  type = 'html', 
-                                 column.labels = c("coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio", 
-                                                   "coefficients(se)", "Odds ratio"),
+                                 column.labels = c("coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR", 
+                                                   "coefficient (se)", "OR"),
                                  model.numbers = FALSE, 
-                                 dep.var.labels = c("operator cert violations", "M&R violations", "Cash on hand"), 
+                                 dep.var.labels = c("Operator cert. violations", "M&R violations", "Cash on hand"), 
                                  ci=c(F,T, F, T, F, T, F, T), 
                                  star.cutoffs = c(0.05, 0.01, 0.001),
+                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
+                                 notes.append = FALSE,
                                  p = c(pvaluesOpcert, p.values2, pvaluesmrviolations, p.values2, pvaluescash, p.values2),
                                  omit.stat = c("ll", "aic"),
-                                 covariate.labels = c("Lim enfran.", "No enfran.",
+                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                       "Population (log)", "Surface water",
-                                                      "Purchased"),
+                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                  omit = "Constant",
@@
 ```{r descriptives for demofraphics, results='asis', echo=FALSE}
-Data_mini <- Data %>% select(enfranchisement_final, MHI_KD, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, TMF_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian)
+Data_mini <- Data %>% select(enfranchisement_final, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, TMF_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian)
 
@@
   summarize(meanPOP = mean(na.omit(POPULATION)),
-            meanMHI_KD = mean(na.omit(MHI_KD)),
             meanMHI = mean(na.omit(MHI)),
@@
 #ANOVA and Tukey HSD
-MHI <- aov(MHI_KD ~ enfranchisement_final, data = Data)
+MHI <- aov(MHI ~ enfranchisement_final, data = Data)
 summary(MHI)
@@
 TukeyMHI_table <- as.data.frame(TukeyMHI$enfranchisement_final)
+
+Data$Source_bi <- NA
+Data$Source_bi <- ifelse(Data$Source == "GW", 1, Data$Source_bi)
+Data$Source_bi <- ifelse(Data$Source == "SW", 0, Data$Source_bi)
+
+GW <- aov(Data$Source_bi ~ Data$enfranchisement_final)
+summary(GW)
+TukeyGW <- TukeyHSD(GW)
+TukeyGW_table <- as.data.frame(TukeyGW$`Data$enfranchisement_final`)
+
+Pop <- aov(Data$POPULATION ~ Data$enfranchisement_final)
+summary(Pop)
+TukeyPop <- TukeyHSD(Pop)
+TukeyPop_table <- as.data.frame(TukeyPop$`Data$enfranchisement_final`)
+
+Rent <- aov(Data$Percent.renter ~ Data$enfranchisement_final)
+summary(Rent)
+TukeyRent <- TukeyHSD(Rent)
+TukeyRent_table <- as.data.frame(TukeyRent$`Data$enfranchisement_final`)
+
+CalE <- aov(Data$CALENVIRO_SCREEN_SCORE ~ Data$enfranchisement_final)
+summary(CalE)
+TukeyCalE <- TukeyHSD(CalE)
+TukeyCalE_table <- as.data.frame(TukeyCalE$`Data$enfranchisement_final`)
+
+White <- aov(Data$Percent.white ~ Data$enfranchisement_final)
+summary(White)
+TukeyWhite <- TukeyHSD(White)
+TukeyWhite_table <- as.data.frame(TukeyWhite$`Data$enfranchisement_final`)
 
@@
 TukeyAA_table <- as.data.frame(TukeyAA$enfranchisement_final)
+
+Lat <- aov(Data$Percent.hispanicorlatino ~ Data$enfranchisement_final)
+summary(Lat)
+TukeyLat <- TukeyHSD(Lat)
+TukeyLat_table <- as.data.frame(TukeyLat$`Data$enfranchisement_final`)
+
+As <- aov(Data$Percent.asian ~ Data$enfranchisement_final)
+summary(As)
+TukeyAs <- TukeyHSD(As)
+TukeyAs_table <- as.data.frame(TukeyAs$`Data$enfranchisement_final`)
+
 ```

### 2024-12-20 Code/Data_compiliation_SES.R

@@ -76,8 +76,8 @@ summary(Data_dem$Percent.renter)
 #Add in county
 SDWIS <- read_csv("Data_raw/CA_SDWIS_allcws_Nov152024.csv")
 SDWIS <- SDWIS[,c(1,6)]
-SDWIS <- SDWIS %>% rename("COUNTY" = "Principal County Served")
-SDWIS <- SDWIS %>% rename("SABL_PWSID" = "Water System No.")
+SDWIS <- SDWIS %>% dplyr::rename("COUNTY" = "Principal County Served")
+SDWIS <- SDWIS %>% dplyr::rename("SABL_PWSID" = "Water System No.")
 Data_geo <- left_join(Data_geo, SDWIS)
 
 #save data_geo to make map later

### 2024-12-20 Code/Data_compiliation_SES_updated2022.R

@@ -1,89 +1,0 @@
-#Data compilation for demographics updated to 2022 five year estimates
-
-#LOAD LIBRARIES
-library(tidyverse)
-library(ggmap)
-library(censusapi)
-library(tidycensus)
-library(sf)
-library(sp)
-
-Demographics_blockg <- get_acs(geography = "block group", variables = c("B03002_001E", "B03002_003E", "B03002_004E", "B03002_005E", "B03002_006E", "B03002_007E", "B03002_012E", "B19013_001E", "B25003_001E", "B25003_002E", "B25003_003E", "B11001_001E"), state = "CA", year =2022, output = "wide", survey = "acs5", geometry = TRUE)
-
-#rename variables
-Demographics_blockg <- Demographics_blockg %>% dplyr::rename(Race.estimate.total = B03002_001E, Race.white.alone = B03002_003E, Race.black.alone = B03002_004E, Race.native.alone = B03002_005E, Race.asian.alone = B03002_006E, Race.PI.alone = B03002_007E, Race.hispanicorlatino = B03002_012E, Median.hh.income = B19013_001E, Tenure.estimate.total = B25003_001E, Tenure.owner = B25003_002E, Tenure.renter = B25003_003E, Households.total = B11001_001E)
-
-#Make median household income into a count by multiplying median income by total households (see https://crd230.github.io/lab3.html)
-Demographics_blockg$Income.aggregatecount <- Demographics_blockg$Median.hh.income*Demographics_blockg$Households.total
-
-#load in boundary polygons for CWS
-PWS_boundary <- st_read("Data_raw/California_Drinking_Water_System_Area_Boundaries/") #New version of SABL downloaded November 15, 2024
-str(PWS_boundary)
-#plot(st_geometry(PWS_boundary)) #takes a while but works!
-
-PWS_boundary$SABL_PWSID <- as.factor(PWS_boundary$SABL_PWSID)
-PWS_boundary$WATER_SYST <- as.factor(PWS_boundary$WATER_SYST)
-PWS_boundary <- PWS_boundary %>% filter(BOUNDARY_T != "Jurisdictional") #remove jurisdictional boundaries to get rid of duplicates
-
-Data_small <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
-Data_small <- Data_small[,c(2,11)]
-
[…]

### 2024-12-20 Code/Figures_PP.R

@@ -64,9 +64,9 @@ Effect_AF_covid <- predictions(covid,
                                newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
 
 
-Effect_AF_MHI$Key <- "Bill ≥ 1.5% of MHI"
-Effect_AF_extreme$Key <- "Bill ≥ 150% of statewide average"
-Effect_AF_covid$Key <- "Did not apply for arrearage relief"
+Effect_AF_MHI$Key <- "Unaffordable bill"
+Effect_AF_extreme$Key <- "Extreme bill"
+Effect_AF_covid$Key <- "No arrearage relief"
 Probabilities_AF <- bind_rows(Effect_AF_MHI, Effect_AF_extreme, Effect_AF_covid)
 Probabilities_AF$Category <- "Affordability"
 
@@ -108,7 +108,7 @@ library(RColorBrewer)
 #Probabilities_MASSIVE <- bind_rows(Effect_WQ_ecoli, Effect_WQ_CEC, Effect_WQ_MCL, Effect_WQ_TT, Effect_AF_MHI, Effect_AF_extreme, Effect_AF_covid, Effect_AF_funding, Effect_AC_sources, Effect_AC_interties, Effect_AC_bottled, Effect_AC_sourcecapacity, Effect_TMF_opcertviolations, Effect_TMF_mrviolations, Effect_TMF_cash, Effect_TMF_operating)
 Probabilities_MASSIVE <- bind_rows(Probabilities_WQ, Probabilities_AF, Probabilities_TMF)
 Probabilities_MASSIVE$Key <- as.factor(Probabilities_MASSIVE$Key)
-Probabilities_MASSIVE$Key <- factor(Probabilities_MASSIVE$Key, levels=c("E. coli", "MCL exceedences", "Treatment Technique violations", "Bill ≥ 150% of statewide average", "Bill ≥ 1.5% of MHI", "Did not apply for arrearage relief", "Operator certification violations", "M&R violations", "<30 days cash on hand"))
+Probabilities_MASSIVE$Key <- factor(Probabilities_MASSIVE$Key, levels=c("E. coli", "MCL exceedences", "Treatment Technique violations", "Extreme bill", "Unaffordable bill", "No arrearage relief", "Operator certification violations", "M&R violations", "<30 days cash on hand"))
 
 plotCombined_MASSIVE <- Probabilities_MASSIVE %>%
   ggplot(aes(x = enfranchisement_final, y = estimate*100)) +

### 2024-12-20 Code/Sensitivity_analysis.R

@@ -142,7 +142,7 @@ plotCombined_alt <- Probabilities_alt %>%
   facet_wrap(~Key, scales = "free", nrow = 1, ncol = 3) +
   cowplot::theme_half_open(); plotCombined_alt
 
-ggsave("Fig2_alt.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
+ggsave("Fig2_alt.jpeg", plotCombined_alt, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
 
 #water quality subindicators figure
 Effect_WQ_ecoli_alt <- predictions(ecoli_alt,
@@ -178,9 +178,9 @@ Effect_AF_covid_alt <- predictions(covid_alt,
                                newdata = datagrid(enfranchisement_alt = c("Full - city", "Full - other", "Limited", "None - MHP", "None - IOU")))
 
 
-Effect_AF_MHI_alt$Key <- "Bill ≥ 1.5% of MHI"
-Effect_AF_extreme_alt$Key <- "Bill ≥ 150% of statewide average"
-Effect_AF_covid_alt$Key <- "Did not apply for arrearage relief"
+Effect_AF_MHI_alt$Key <- "Unaffordable bill"
+Effect_AF_extreme_alt$Key <- "Extreme bill"
+Effect_AF_covid_alt$Key <- "No arrearage relief"
 Probabilities_AF_alt <- bind_rows(Effect_AF_MHI_alt, Effect_AF_extreme_alt, Effect_AF_covid_alt)
 Probabilities_AF_alt$Category <- "Affordability"
 
@@ -195,7 +195,7 @@ plotCombined_AF_alt <- Probabilities_AF_alt %>%
   cowplot::theme_half_open() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)); plotCombined_AF_alt
 
-ggsave("FigSIAF_alt.jpeg", plotCombined_AF_alt, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
+ggsave("FigSIAF_alt.jpeg", plotCombined_AF_alt, path = "Figures/", height = 6, width = 9, units = "in", dpi = 720)
 
 #TMF subindicators figure
 Effect_TMF_opcertviolations_alt <- predictions(Opcert_violations_alt,

### 2025-07-31 Code/Paper_analysisandtables_R&R.Rmd
Clean out repository for manuscript publication getting rid of old versions of data and code

---
title: "Paper_analysisandtables"
author: "Kristin Dobbin"
date: "2024-09-18"
output: 
  html_document: default
  pdf_document: default
knit: (function(rmdfile, ...) { rmarkdown::render(rmdfile, output_dir=here::here("Docs"))})
---

```{r setup, include=FALSE, echo=FALSE}
#MARKDOWN SETTINGS
#knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Documents/Projects/R_Projects/Votingrights_systemperformance", warning=FALSE, message=FALSE, include=FALSE)

#LOAD LIBRARIES
library(tidyverse)
library(vtable)
library(sjPlot)
library(stargazer)

#LOAD AND READY DATA
Data <- read.csv(here::here("Data_processed/Systemperformancedata_RR.csv"))
Data <- Data[,-1]
Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
Data$Final_inst_update <- as.factor(Data$Final_inst_update)

#correct data types, levels and reference factors
Data[,c(1:11,16, 19:45, 47:61)] <- lapply(Data[,c(1:11,16, 19:45, 47:61)], factor)
levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")

Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
Data$FAILING <- as.factor(Data$FAILING)

Data$Source <- as.factor(Data$Source)
Data$public <- as.factor(Data$public)

#re-order factor levels for enfranchisement to be full, limited, none
Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))

Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")

#Make self produced water the reference (compared to purchased)
Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")

Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI <- relevel(Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI, ref = "NONE")
Data$ABESENCE_OF_INTERTIES_RISK_LEVEL <- relevel(Data$ABESENCE_OF_INTERTIES_RISK_LEVEL, ref = "NONE")
Data$BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL <- relevel(Data$BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL, ref = "NONE")
Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL <- relevel(Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL, ref = "NONE")
Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI <- relevel(Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI, ref = "2")
Data$EXTREME_WATER_BILL_RISK_LEVEL_BI <- relevel(Data$EXTREME_WATER_BILL_RISK_LEVEL_BI, ref = "2")
Data$FUNDING_any_failingoratriskonly <- relevel(Data$FUNDING_any_failingoratriskonly, ref = "No")
Data$OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL <- relevel(Data$OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL <- relevel(Data$MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$DAYS_CASH_ON_HAND_RISK_LEVEL <- relevel(Data$DAYS_CASH_ON_HAND_RISK_LEVEL, ref = "NONE")
Data$OPERATING_RATIO_RISK_LEVEL <- relevel(Data$OPERATING_RATIO_RISK_LEVEL, ref = "NONE")

#add in demographic data
Demographics <- read.csv(here::here("Data_processed/Systemdemographicdata.csv"))
Demographics <- Demographics[,-c(1,3)]
Demographics$PWSID <- as.factor(Demographics$PWSID)

Data <- left_join(Data, Demographics, by = "PWSID")

#Water quality
WQ1 <- glm(WQ_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(WQ1)

##WQ sub components
ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(ecoli)

TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(TT)

MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(MCL)

#Affordability 
Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Affordability1)

## Affordability sub components
MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(MHI)

extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(extreme)

covid <- glm(didnotapplycovid ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(covid)

# Accessibility
Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Accessibility1)

## Accessibility sub components
Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Single_source)

Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Sourcecapacity_violations)

Service_dis <- glm(Service_disruptions ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Service_dis)

```

This Rmd created html versions of the model results tables using stargazer. The models script must be run prior to using this Rmd

```{r table composites, results='asis', echo=FALSE}
#Composites
WQ1OR <- WQ1
WQ1OR$coefficients <- exp(WQ1OR$coefficients)
p.valuesWQ1 <- list(summary(WQ1)$coefficients[,4])

Affordability1OR <- Affordability1
Affordability1OR$coefficients <- exp(Affordability1OR$coefficients)
pvaluesAffordability1 <- list(summary(Affordability1)$coefficients[,4])

Accessibility1OR <- Accessibility1
Accessibility1OR$coefficients <- exp(Accessibility1OR$coefficients)
pvaluesAccessibility1 <- list(summary(Accessibility1)$coefficients[,4])

p.values2 <- list(c(1,1,1,1,1,1))

compositemodellist <- list(WQ1, WQ1OR, Affordability1, Affordability1OR,
                          Accessibility1, Accessibility1OR)

Compositestable <- capture.output(stargazer(compositemodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("Water quality", "Affordability", 
                                                    "Accessibility"), 
                                 ci=c(F,T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(p.valuesWQ1, p.values2, pvaluesAffordability1, p.values2,
                                       pvaluesAccessibility1, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", Compositestable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table WQ, results='asis', echo=FALSE}
#WQ indicators
ecoliOR <- ecoli
ecoliOR$coefficients <- exp(ecoliOR$coefficients)
pvaluesecoli <- list(summary(ecoli)$coefficients[,4])

TTOR <- TT
TTOR$coefficients <- exp(TTOR$coefficients)
pvaluesTT <- list(summary(TT)$coefficients[,4])

MCLOR <- MCL
MCLOR$coefficients <- exp(MCLOR$coefficients)
pvaluesMCL <- list(summary(MCL)$coefficients[,4])

WQmodellist <- list(ecoli, ecoliOR, TT, TTOR, MCL, MCLOR)

WQtable <- capture.output(stargazer(WQmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("E. coli", "Treatment technique violations", "MCL violations"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(pvaluesecoli, p.values2, pvaluesTT, p.values2,
                                       pvaluesMCL, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", WQtable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table affordability, results='asis', echo=FALSE}
#affordability indicators
MHIOR <- MHI
MHIOR$coefficients <- exp(MHIOR$coefficients)
pvaluesMHI <- list(summary(MHI)$coefficients[,4])

extremeOR <- extreme
extremeOR$coefficients <- exp(extremeOR$coefficients)
pvaluesextreme <- list(summary(extreme)$coefficients[,4])

covidOR <- covid
covidOR$coefficients <- exp(covidOR$coefficients)
pvaluescovid <- list(summary(covid)$coefficients[,4])

AFmodellist <- list(MHI, MHIOR, extreme, extremeOR, covid, covidOR)

AFtable <- capture.output(stargazer(AFmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("Unaffordable bill", "Extreme bill",
                                                    "No arrearage relief"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(pvaluesMHI, p.values2, pvaluesextreme, p.values2, pvaluescovid, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", AFtable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table Accessibility, results='asis', echo=FALSE}
#Accessibility indicators
Single_sourceOR <- Single_source
Single_sourceOR$coefficients <- exp(Single_sourceOR$coefficients)
pvaluesSinglesource <- list(summary(Single_source)$coefficients[,4])

Sourcecapacity_violationsOR <- Sourcecapacity_violations
Sourcecapacity_violationsOR$coefficients <- exp(Sourcecapacity_violationsOR$coefficients)
pvaluesSourcecapacity_violations <- list(summary(Sourcecapacity_violations)$coefficients[,4])

Service_disOR <- Service_dis
Service_disOR$coefficients <- exp(Service_disOR$coefficients)
pvaluesService_dis <- list(summary(Service_dis)$coefficients[,4])

Accessmodellist <- list(Single_source, Single_sourceOR, Sourcecapacity_violations, Sourcecapacity_violationsOR, Service_dis, Service_disOR)

Accesstable <- capture.output(stargazer(Accessmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("Single source", "Source capacity violations", "Service disruptions"), 
                                 ci=c(F,T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(pvaluesSinglesource, p.values2, pvaluesSourcecapacity_violations, p.values2, pvaluesService_dis, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", Accesstable), collapse = "\n"), "\n")
```

<br>
<br>

```{r descriptives for demofraphics, results='asis', echo=FALSE}

Data$Source_bi <- NA
Data$Source_bi <- ifelse(Data$Source == "GW", 1, Data$Source_bi)
Data$Source_bi <- ifelse(Data$Source == "SW", 0, Data$Source_bi)

Data_mini <- Data %>% select(enfranchisement_final, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian, Source_bi)

Descriptives<- Data_mini %>%
  group_by(enfranchisement_final) %>% 
  summarize(meanPOP = mean(na.omit(POPULATION)),
            meanMHI = mean(na.omit(MHI)),
            meanPercRent = mean(na.omit(Percent.renter)),
            meanPercWhite = mean(na.omit(Percent.white)),
            meanCES = mean(na.omit(CALENVIRO_SCREEN_SCORE)),
            medianpop = median(na.omit(POPULATION)))

SIdescriptives <- Data_mini %>% 
  group_by(enfranchisement_final) %>% 
  select_if(is.numeric) %>% 
  gather(Variable, Value, -enfranchisement_final) %>% 
  group_by(Variable, enfranchisement_final) %>% 
  summarise(min = min(na.omit(Value)), 
            median = median(na.omit(Value)),
            mean = mean(na.omit(Value)), 
            max = max(na.omit(Value)),
            sd = sd(na.omit(Value))) %>% 
 ungroup()

SIdescriptives <- SIdescriptives %>% mutate_if(is.numeric, round, digits=3)
```

```{r dem sig tests, results='asis', echo=FALSE}
#ANOVA and Tukey HSD
MHI <- aov(MHI ~ enfranchisement_final, data = Data)
summary(MHI)
TukeyMHI <- TukeyHSD(MHI)
TukeyMHI_table <- as.data.frame(TukeyMHI$enfranchisement_final)

GW <- aov(Data$Source_bi ~ Data$enfranchisement_final)
summary(GW)
TukeyGW <- TukeyHSD(GW)
TukeyGW_table <- as.data.frame(TukeyGW$`Data$enfranchisement_final`)

Pop <- aov(Data$POPULATION ~ Data$enfranchisement_final)
summary(Pop)
TukeyPop <- TukeyHSD(Pop)
TukeyPop_table <- as.data.frame(TukeyPop$`Data$enfranchisement_final`)

Rent <- aov(Data$Percent.renter ~ Data$enfranchisement_final)
summary(Rent)
TukeyRent <- TukeyHSD(Rent)
TukeyRent_table <- as.data.frame(TukeyRent$`Data$enfranchisement_final`)

CalE <- aov(Data$CALENVIRO_SCREEN_SCORE ~ Data$enfranchisement_final)
summary(CalE)
TukeyCalE <- TukeyHSD(CalE)
TukeyCalE_table <- as.data.frame(TukeyCalE$`Data$enfranchisement_final`)

White <- aov(Data$Percent.white ~ Data$enfranchisement_final)
summary(White)
TukeyWhite <- TukeyHSD(White)
TukeyWhite_table <- as.data.frame(TukeyWhite$`Data$enfranchisement_final`)

AA <- aov(Percent.black ~ enfranchisement_final, data = Data)
summary(AA)
TukeyAA <- TukeyHSD(AA)
TukeyAA_table <- as.data.frame(TukeyAA$enfranchisement_final)

Lat <- aov(Data$Percent.hispanicorlatino ~ Data$enfranchisement_final)
summary(Lat)
TukeyLat <- TukeyHSD(Lat)
TukeyLat_table <- as.data.frame(TukeyLat$`Data$enfranchisement_final`)

As <- aov(Data$Percent.asian ~ Data$enfranchisement_final)
summary(As)
TukeyAs <- TukeyHSD(As)
TukeyAs_table <- as.data.frame(TukeyAs$`Data$enfranchisement_final`)

```

### 2025-07-31 Code/sensitivity_tests_march2025_tables_profit.Rmd
Clean out repository for manuscript publication getting rid of old versions of data and code

--- 2025-07-31 Code/Paper_analysisandtables_R&R.Rmd
@@
 ---
-title: "Paper_analysisandtables"
+title: "sensitivity_tests_march2025_tables_profit"
 author: "Kristin Dobbin"
-date: "2024-09-18"
+date: "2025-04-14"
 output: 
@@
 
+Data$profit <- ifelse(Data$Final_inst_update == "Investor Owned Utility" | 
+                        Data$Final_inst_update == "Mobile Home Park", "Yes", "No")
+
+Data$profit <- as.factor(Data$profit)
+summary(Data$profit)
+Data$profit <- relevel(Data$profit, ref = "No")
+
 #Water quality
-WQ1 <- glm(WQ_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+
+WQ1 <- glm(WQ_combined_any ~  profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(WQ1)
@@
 ##WQ sub components
-ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+
+ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(ecoli)
 
-TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(TT)
 
-MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MCL)
@@
 #Affordability
-Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+
+Affordability1 <- glm(Affordability_combined_any ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Affordability1)
@@
 ## Affordability sub components
-MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+
+MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MHI)
 
-extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(extreme)
 
-covid <- glm(didnotapplycovid ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+covid <- glm(didnotapplycovid ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(covid)
@@
 # Accessibility
-Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+
+Accessibility1 <- glm(Accessibility_combined_any ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Accessibility1)
@@
 ## Accessibility sub components
-Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+
+Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Single_source)
 
-Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Sourcecapacity_violations)
 
-Service_dis <- glm(Service_disruptions ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Service_dis <- glm(Service_disruptions ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Service_dis)
@@
                                  omit.stat = c("ll"),
-                                 covariate.labels = c("Limited enfran.", "No enfran.",
+                                 covariate.labels = c("For profit system",
                                                       "Population", "Surface water",
@@
                                  omit.stat = c("ll"),
-                                 covariate.labels = c("Limited enfran.", "No enfran.",
+                                 covariate.labels = c("For profit system", 
                                                       "Population", "Surface water",
@@
 ```
-
-<br>
-<br>
-
-```{r descriptives for demofraphics, results='asis', echo=FALSE}
-
-Data$Source_bi <- NA
-Data$Source_bi <- ifelse(Data$Source == "GW", 1, Data$Source_bi)
-Data$Source_bi <- ifelse(Data$Source == "SW", 0, Data$Source_bi)
-
-Data_mini <- Data %>% select(enfranchisement_final, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian, Source_bi)
-
-Descriptives<- Data_mini %>%
-  group_by(enfranchisement_final) %>% 
-  summarize(meanPOP = mean(na.omit(POPULATION)),
-            meanMHI = mean(na.omit(MHI)),
-            meanPercRent = mean(na.omit(Percent.renter)),
-            meanPercWhite = mean(na.omit(Percent.white)),
-            meanCES = mean(na.omit(CALENVIRO_SCREEN_SCORE)),
-            medianpop = median(na.omit(POPULATION)))
-
-SIdescriptives <- Data_mini %>% 
-  group_by(enfranchisement_final) %>% 
-  select_if(is.numeric) %>% 
-  gather(Variable, Value, -enfranchisement_final) %>% 
-  group_by(Variable, enfranchisement_final) %>% 
-  summarise(min = min(na.omit(Value)), 
-            median = median(na.omit(Value)),
-            mean = mean(na.omit(Value)), 
-            max = max(na.omit(Value)),
-            sd = sd(na.omit(Value))) %>% 
- ungroup()
-
-SIdescriptives <- SIdescriptives %>% mutate_if(is.numeric, round, digits=3)
-```
-
-```{r dem sig tests, results='asis', echo=FALSE}
-#ANOVA and Tukey HSD
-MHI <- aov(MHI ~ enfranchisement_final, data = Data)
-summary(MHI)
-TukeyMHI <- TukeyHSD(MHI)
-TukeyMHI_table <- as.data.frame(TukeyMHI$enfranchisement_final)
-
-GW <- aov(Data$Source_bi ~ Data$enfranchisement_final)
-summary(GW)
-TukeyGW <- TukeyHSD(GW)
-TukeyGW_table <- as.data.frame(TukeyGW$`Data$enfranchisement_final`)
-
-Pop <- aov(Data$POPULATION ~ Data$enfranchisement_final)
-summary(Pop)
-TukeyPop <- TukeyHSD(Pop)
-TukeyPop_table <- as.data.frame(TukeyPop$`Data$enfranchisement_final`)
-
-Rent <- aov(Data$Percent.renter ~ Data$enfranchisement_final)
-summary(Rent)
-TukeyRent <- TukeyHSD(Rent)
-TukeyRent_table <- as.data.frame(TukeyRent$`Data$enfranchisement_final`)
-
-CalE <- aov(Data$CALENVIRO_SCREEN_SCORE ~ Data$enfranchisement_final)
-summary(CalE)
-TukeyCalE <- TukeyHSD(CalE)
-TukeyCalE_table <- as.data.frame(TukeyCalE$`Data$enfranchisement_final`)
-
-White <- aov(Data$Percent.white ~ Data$enfranchisement_final)
-summary(White)
-TukeyWhite <- TukeyHSD(White)
-TukeyWhite_table <- as.data.frame(TukeyWhite$`Data$enfranchisement_final`)
-
-AA <- aov(Percent.black ~ enfranchisement_final, data = Data)
-summary(AA)
-TukeyAA <- TukeyHSD(AA)
-TukeyAA_table <- as.data.frame(TukeyAA$enfranchisement_final)
-
-Lat <- aov(Data$Percent.hispanicorlatino ~ Data$enfranchisement_final)
-summary(Lat)
-TukeyLat <- TukeyHSD(Lat)
-TukeyLat_table <- as.data.frame(TukeyLat$`Data$enfranchisement_final`)
-
-As <- aov(Data$Percent.asian ~ Data$enfranchisement_final)
-summary(As)
-TukeyAs <- TukeyHSD(As)
-TukeyAs_table <- as.data.frame(TukeyAs$`Data$enfranchisement_final`)
-
-```

### 2025-07-31 Code/sensitivity_tests_march2025_tables_public.Rmd
Clean out repository for manuscript publication getting rid of old versions of data and code

--- 2025-07-31 Code/sensitivity_tests_march2025_tables_profit.Rmd
@@
 ---
-title: "sensitivity_tests_march2025_tables_profit"
+title: "sensitivity_tests_march2025_tables"
 author: "Kristin Dobbin"
@@
 
-Data$profit <- ifelse(Data$Final_inst_update == "Investor Owned Utility" | 
-                        Data$Final_inst_update == "Mobile Home Park", "Yes", "No")
-
-Data$profit <- as.factor(Data$profit)
-summary(Data$profit)
-Data$profit <- relevel(Data$profit, ref = "No")
+#subbing enfranchisement for public private 
+Data$public <- relevel(Data$public, ref = "Yes")
 
@@
 
-WQ1 <- glm(WQ_combined_any ~  profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+WQ1 <- glm(WQ_combined_any ~  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(WQ1)
@@
 
-ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(ecoli)
 
-TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(TT)
 
-MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MCL)
@@
 
-Affordability1 <- glm(Affordability_combined_any ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Affordability1 <- glm(Affordability_combined_any ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Affordability1)
@@
 
-MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MHI)
 
-extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(extreme)
 
-covid <- glm(didnotapplycovid ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+covid <- glm(didnotapplycovid ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(covid)
@@
 
-Accessibility1 <- glm(Accessibility_combined_any ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Accessibility1 <- glm(Accessibility_combined_any ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Accessibility1)
@@
 
-Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Single_source)
 
-Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Sourcecapacity_violations)
 
-Service_dis <- glm(Service_disruptions ~ profit + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Service_dis <- glm(Service_disruptions ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Service_dis)
@@
                                  omit.stat = c("ll"),
-                                 covariate.labels = c("For profit system",
+                                 covariate.labels = c("Ownership - private",
                                                       "Population", "Surface water",
@@
                                  omit.stat = c("ll"),
-                                 covariate.labels = c("For profit system", 
+                                 covariate.labels = c("Ownership - private", 
                                                       "Population", "Surface water",

### 2025-07-31 Code/sensitivity_tests_march2025_tables_specialization.Rmd
Clean out repository for manuscript publication getting rid of old versions of data and code

--- 2025-07-31 Code/sensitivity_tests_march2025_tables_public.Rmd
@@
 ---
-title: "sensitivity_tests_march2025_tables"
+title: "sensitivity_tests_march2025_tables_specialization"
 author: "Kristin Dobbin"
@@
 
-#subbing enfranchisement for public private 
-Data$public <- relevel(Data$public, ref = "Yes")
+Data$specialized <- ifelse(Data$Final_inst_update == "City" | 
+                          Data$Final_inst_update == "County Waterworks District" |
+                          Data$Final_inst_update == "Maintenance District" |
+                          Data$Final_inst_update == "Mobile Home Park", "No", "Yes")
+
+Data <- Data %>% mutate(specialized = case_when(PWSID == "CA2800526" ~ "No", .default = specialized), 
+                        specialized = case_when(PWSID == "CA2810013" ~ "No", .default = specialized)) #NAPA RIDs are subsidary districts managed by county (along with Glenn county) everywhere else they are independent special districts
+
+Data$specialized <- as.factor(Data$specialized)
+summary(Data$specialized)
+
+Data$specialized <- relevel(Data$specialized, ref = "No")
 
@@
 
-WQ1 <- glm(WQ_combined_any ~  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+WQ1 <- glm(WQ_combined_any ~  enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(WQ1)
@@
 
-ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(ecoli)
 
-TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(TT)
 
-MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MCL)
@@
 
-Affordability1 <- glm(Affordability_combined_any ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Affordability1)
@@
 
-MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MHI)
 
-extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(extreme)
 
-covid <- glm(didnotapplycovid ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+covid <- glm(didnotapplycovid ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(covid)
@@
 
-Accessibility1 <- glm(Accessibility_combined_any ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Accessibility1)
@@
 
-Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Single_source)
 
-Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Sourcecapacity_violations)
 
-Service_dis <- glm(Service_disruptions ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Service_dis <- glm(Service_disruptions ~ enfranchisement_final + specialized + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Service_dis)
@@
                                  omit.stat = c("ll"),
-                                 covariate.labels = c("Ownership - private",
+                                 covariate.labels = c("Limited enfran.", "No enfran.", "Specialized - yes",
                                                       "Population", "Surface water",
@@
                                  omit.stat = c("ll"),
-                                 covariate.labels = c("Ownership - private", 
+                                 covariate.labels = c("Limited enfran.", "No enfran.", "Specialized - yes",
                                                       "Population", "Surface water",

### 2025-07-31 Code/Data_compilation_systemperformance.R

@@ -1,348 +1,0 @@
-#Compile base data
-
-#load libraries
-library(tidyverse)
-library(plyr)
-
-# Create sample and governance variables ----------------------------------------------------
-
-###Load data and initial manipulations
-CWSgov <- read.csv(here::here("Data_raw/Dobbin_Fencl_McBride_2023_CA_CWS_institutional_type_update.csv"), header=T, na.strings=c("","NA"))
-CWSgov$Primacy_FINAL <- as.factor(CWSgov$Primacy_FINAL)
-CWSgov <- CWSgov %>% filter(Primacy_FINAL != "EPA Region 9") #get rid of EPA regulated Tribal water systems
-CWSgov <- CWSgov %>% filter(Final_inst_update != "Tribal Government") #Get rid of three remaining systems that are Tribal but state regulated
-CWSgov$Primacy_FINAL <- droplevels(CWSgov$Primacy_FINAL)
-CWSgov$PWSID <- as.factor(CWSgov$PWSID)
-CWSgov$Final_inst_update <- as.factor(CWSgov$Final_inst_update)
-CWSgov$Inst_Subtype  <- as.factor(CWSgov$Inst_Subtype)
-
-### create new variables for ancillary/non ancillary, public/private and customer enfranchisement
-
-## To focus on residential systems where residents are served and could have a role in governance/decision-making, identify which systems are "residential" this is mostly synonymous to non-ancillary except with this variable maintaining Mobile Home Parks which are generally considered to be ancillary and yet in other ways these systems are more similar to non-ancillary systems in that they serve a stable residential base. Moreover, California has seen significant organizing among Mobile Home Park communities in recent years around drinking water issues. 
-CWSgov$Residential <- ifelse(CWSgov$Final_inst_update == "Community Services District" |
-                               CWSgov$Final_inst_update == "County Service Area" | 
-                               CWSgov$Final_inst_update == "Municipal Water District" | 
-                               CWSgov$Final_inst_update == "City" | 
-                               CWSgov$Final_inst_update == "County Sanitation District"| 
-                               CWSgov$Final_inst_update == "County Water District"| 
-                               CWSgov$Final_inst_update == "County Waterworks District" |
-                               CWSgov$Final_inst_update == "Irrigation District"| 
-                               CWSgov$Final_inst_update == "Maintenance District"|
-                               CWSgov$Final_inst_update == "Municipal Utility District"|
[…]

### 2025-07-31 Code/Data_compilation_systemperformance_R&R.R

@@ -1,0 +1,374 @@
+#Compile base data
+#R&R revised to follow tenants of the HR2W for analysis structure
+
+#load libraries
+library(tidyverse)
+library(plyr)
+
+# Create sample and governance variables ----------------------------------------------------
+
+###Load data and initial manipulations
+CWSgov <- read.csv(here::here("Data_raw/Dobbin_Fencl_McBride_2023_CA_CWS_institutional_type_update.csv"), header=T, na.strings=c("","NA"))
+CWSgov$Primacy_FINAL <- as.factor(CWSgov$Primacy_FINAL)
+CWSgov <- CWSgov %>% filter(Primacy_FINAL != "EPA Region 9") #get rid of EPA regulated Tribal water systems
+CWSgov <- CWSgov %>% filter(Final_inst_update != "Tribal Government") #Get rid of three remaining systems that are Tribal but state regulated
+CWSgov$Primacy_FINAL <- droplevels(CWSgov$Primacy_FINAL)
+CWSgov$PWSID <- as.factor(CWSgov$PWSID)
+CWSgov$Final_inst_update <- as.factor(CWSgov$Final_inst_update)
+CWSgov$Inst_Subtype  <- as.factor(CWSgov$Inst_Subtype)
+
+### create new variables for ancillary/non ancillary, public/private and residential enfranchisement
+
+## To focus on residential systems where residents are served and could have a role in governance/decision-making, identify which systems are "residential" this is mostly synonymous to non-ancillary except with this variable maintaining Mobile Home Parks which are generally considered to be ancillary and yet in other ways these systems are more similar to non-ancillary systems in that they serve a stable residential base. Moreover, California has seen significant organizing among Mobile Home Park communities in recent years around drinking water issues. 
+CWSgov$Residential <- ifelse(CWSgov$Final_inst_update == "Community Services District" |
+                               CWSgov$Final_inst_update == "County Service Area" | 
+                               CWSgov$Final_inst_update == "Municipal Water District" | 
+                               CWSgov$Final_inst_update == "City" | 
+                               CWSgov$Final_inst_update == "County Sanitation District"| 
+                               CWSgov$Final_inst_update == "County Water District"| 
+                               CWSgov$Final_inst_update == "County Waterworks District" |
+                               CWSgov$Final_inst_update == "Irrigation District"| 
+                               CWSgov$Final_inst_update == "Maintenance District"|
+                               CWSgov$Final_inst_update == "Municipal Utility District"|
[…]

### 2025-07-31 Code/Figures_PP.R

@@ -1,131 +1,0 @@
-#Predicted probabilities figures for GLMs. NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Models.R script
-library(marginaleffects)
-
-#pull out probabilities in confidence intervals
-Effect_WQ <- predictions(WQ1,
-                         newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
-Effect_Affordability <- predictions(Affordability1,
-                                    newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
-Effect_TMF <- predictions(TMF1,
-                          newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
-
-#add key and combine data sets
-Effect_WQ$Key <- "Water quality risk"
-Effect_Affordability$Key <- "Affordability risk"
-Effect_TMF$Key <- "TMF capacity risk"
-Probabilities <- bind_rows(Effect_WQ, Effect_TMF, Effect_Affordability)
-Probabilities$Key <- factor(Probabilities$Key, levels = c("Water quality risk", "TMF capacity risk", "Affordability risk"))
-
-
-#make figure faceting with key
-plotCombined <- Probabilities %>%
-  ggplot(aes(x = enfranchisement_final, y = estimate*100)) +
-  geom_errorbar(aes(ymin = conf.low*100, ymax = conf.high*100, color = enfranchisement_final), width = 0.2, show.legend = FALSE) +
-  geom_point(aes(color = enfranchisement_final), show.legend = FALSE) +
-  xlab(" ") +
-  ylab(" ") +
-  scale_color_brewer(palette = "Dark2") +
-  facet_wrap(~Key, scales = "free", nrow = 1, ncol = 3) +
-  cowplot::theme_half_open(); plotCombined
-
-ggsave("Fig2.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
-
-#water quality subindicators figure
-Effect_WQ_ecoli <- predictions(ecoli,
-                               newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
-Effect_WQ_TT <- predictions(TT,
-                            newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
[…]

### 2025-07-31 Code/Figures_PP_RR.R

@@ -1,0 +1,131 @@
+#Predicted probabilities figures for GLMs. NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Models.R script
+library(marginaleffects)
+
+#pull out probabilities in confidence intervals
+Effect_WQ <- predictions(WQ1,
+                         newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+Effect_Affordability <- predictions(Affordability1,
+                                    newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+Effect_Access <- predictions(Accessibility1,
+                          newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+
+#add key and combine data sets
+Effect_WQ$Key <- "Water quality risk"
+Effect_Affordability$Key <- "Affordability risk"
+Effect_Access$Key <- "Accessibility risk"
+Probabilities <- bind_rows(Effect_WQ, Effect_Access, Effect_Affordability)
+Probabilities$Key <- factor(Probabilities$Key, levels = c("Water quality risk", "Accessibility risk", "Affordability risk"))
+
+
+#make figure faceting with key
+plotCombined <- Probabilities %>%
+  ggplot(aes(x = enfranchisement_final, y = estimate*100)) +
+  geom_errorbar(aes(ymin = conf.low*100, ymax = conf.high*100, color = enfranchisement_final), width = 0.2, show.legend = FALSE) +
+  geom_point(aes(color = enfranchisement_final), show.legend = FALSE) +
+  xlab(" ") +
+  ylab(" ") +
+  scale_color_brewer(palette = "Dark2") +
+  facet_wrap(~Key, scales = "free", nrow = 1, ncol = 3) +
+  cowplot::theme_half_open(); plotCombined
+
+ggsave("Fig2_RR.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
+
+#water quality subindicators figure
+Effect_WQ_ecoli <- predictions(ecoli,
+                               newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
+Effect_WQ_TT <- predictions(TT,
+                            newdata = datagrid(enfranchisement_final = c("Full", "Limited", "None")))
[…]

### 2025-07-31 Code/Models.R

@@ -1,101 +1,0 @@
-# Models (This code should match Paper_analysisandtables.Rmd)
-
-#LOAD LIBRARIES
-library(tidyverse)
-library(vtable)
-library(sjPlot)
-
-#LOAD AND READY DATA
-Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
-Data <- Data[,-1]
-Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
-Data$Final_inst_update <- as.factor(Data$Final_inst_update)
-
-#correct data types, levels and reference factors
-Data[,c(1:11,16, 19:44, 46:61)] <- lapply(Data[,c(1:11,16, 19:44, 46:61)], factor)
-levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")
-
-Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
-Data$FAILING <- as.factor(Data$FAILING)
-
-Data$Source <- as.factor(Data$Source)
-Data$public <- as.factor(Data$public)
-
-#re-order factor levels for enfranchisement to be full, limited, none
-Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))
-
-Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
-Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")
-
-#Make self produced water the reference (compared to purchased)
-Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")
-
-Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
-Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
-Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
[…]

### 2025-07-31 Code/Models_RR.R

@@ -1,0 +1,109 @@
+# Models (This code should match Paper_analysisandtables_RR.Rmd)
+#Revised for R&R
+
+#LOAD LIBRARIES
+library(tidyverse)
+library(vtable)
+library(sjPlot)
+
+#LOAD AND READY DATA
+Data <- read.csv(here::here("Data_processed/Systemperformancedata_RR.csv"))
+Data <- Data[,-1]
+Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
+Data$Final_inst_update <- as.factor(Data$Final_inst_update)
+
+#correct data types, levels and reference factors
+Data[,c(1:11,16, 19:45, 47:61)] <- lapply(Data[,c(1:11,16, 19:45, 47:61)], factor)
+levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")
+
+Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
+Data$FAILING <- as.factor(Data$FAILING)
+
+Data$Source <- as.factor(Data$Source)
+Data$public <- as.factor(Data$public)
+
+#re-order factor levels for enfranchisement to be full, limited, none
+Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))
+
+Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
+Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")
+
+#Make self produced water the reference (compared to purchased)
+Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")
+
+Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
+Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
+Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
[…]

### 2025-07-31 Code/Sensitivity_analysis.R

@@ -1,243 +1,0 @@
-#Sensitivity analysis - breaking out enfranchisement types more
-
-
-#LOAD LIBRARIES
-library(tidyverse)
-library(vtable)
-library(sjPlot)
-
-#LOAD AND READY DATA
-Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
-Data <- Data[,-1]
-Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
-Data$Final_inst_update <- as.factor(Data$Final_inst_update)
-
-#correct data types, levels and reference factors
-Data[,c(1:11,16, 19:44, 46:61)] <- lapply(Data[,c(1:11,16, 19:44, 46:61)], factor)
-levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")
-
-Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
-Data$FAILING <- as.factor(Data$FAILING)
-
-Data$Source <- as.factor(Data$Source)
-Data$public <- as.factor(Data$public)
-
-#re-order factor levels for enfranchisement to be full, limited, none
-Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))
-
-Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
-Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")
-
-#Make self produced water the reference (compared to purchased)
-Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")
-
-Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
-Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
-Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
[…]

### 2025-07-31 Code/sensitivity_tests_march2025.R

@@ -1,0 +1,455 @@
+#Sensitivity tests March 2025
+#MUST RUN models script before this one
+
+#subbing enfranchisement for public private 
+Data$public <- relevel(Data$public, ref = "Yes")
+
+#Water quality
+summary(WQ1)
+
+WQ1P <- glm(WQ_combined_any ~  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(WQ1P)
+VIF(WQ1P)
+
+WQ1P2 <- glm(WQ_combined_any ~ enfranchisement_final +  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(WQ1P2)
+VIF(WQ1P2)
+
+##WQ sub components
+
+summary(ecoli)
+VIF(ecoli)
+
+ecoliP <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(ecoliP)
+
+ecoliP2 <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final +  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(ecoliP2)
+VIF(ecoliP2)
+
+summary(TT)
+
+TTP <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(TTP)
+
+TTP2 <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(TTP2)
+
+
+summary(MCL)
+
+MCLP <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(MCLP)
+
[…]

### 2025-07-31 Sensitivitytest_MHP.R

@@ -1,124 +1,0 @@
-# Sensitivity analysis making MHPs their own category
-
-#LOAD LIBRARIES
-library(tidyverse)
-library(vtable)
-library(sjPlot)
-
-#LOAD AND READY DATA
-Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
-Data <- Data[,-1]
-Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
-Data$Final_inst_update <- as.factor(Data$Final_inst_update)
-
-#correct data types, levels and reference factors
-Data[,c(1:11,16, 19:44, 46:61)] <- lapply(Data[,c(1:11,16, 19:44, 46:61)], factor)
-levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")
-
-Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
-Data$FAILING <- as.factor(Data$FAILING)
-
-Data$Source <- as.factor(Data$Source)
-Data$public <- as.factor(Data$public)
-
-#re-order factor levels for enfranchisement to be full, limited, none
-Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))
-
-Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
-Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")
-
-#Make self produced water the reference (compared to purchased)
-Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")
-
-Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
-Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
-Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
[…]

### 2025-07-31 Code/Data_compilation_systemperformance.R
Additional repository cleaning, update readme

@@ -1,5 +1,4 @@
-#Compile base data
-#R&R revised to follow tenants of the HR2W for analysis structure
+#Compile base data on water system performance and enfranchisement
 
 #load libraries
 library(tidyverse)

### 2025-07-31 Code/Data_compiliation_SES.R

@@ -1,4 +1,4 @@
-#Data compilation for demographics
+#Data compilation for demographics for water systems
 
 #LOAD LIBRARIES
 library(tidyverse)

### 2025-07-31 Code/Figures_Map.R

@@ -1,4 +1,5 @@
-#Map systems by enfranchisement type. NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Data_compillation_SES.R script
+#Map systems by enfranchisement type.
+## NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Data_compilation_SES.R script
 
 #Load libraries
 library(tidyverse)

### 2025-07-31 Code/Figures_PP.R

@@ -1,4 +1,6 @@
-#Predicted probabilities figures for GLMs. NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Models.R script
+#Predicted probabilities figures for GLMs.
+## NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Models.R script
+
 library(marginaleffects)
 
 #pull out probabilities in confidence intervals

### 2025-07-31 Code/Models.R

@@ -1,5 +1,6 @@
-# Models (This code should match Paper_analysisandtables_RR.Rmd)
-#Revised for R&R
+# Models
+## Script includes creation of the "Fulldataset.csv" file which is the mater data for this project
+## This code should match Paper_analysisandtables.Rmd which is used to generate tables
 
 #LOAD LIBRARIES
 library(tidyverse)

### 2025-07-31 Code/sensitivity_tests_march2025.R

@@ -1,5 +1,5 @@
-#Sensitivity tests March 2025
-#MUST RUN models script before this one
+#Sensitivity tests conducted March 2025
+## MUST RUN models script before this one
 
 #subbing enfranchisement for public private 
 Data$public <- relevel(Data$public, ref = "Yes")

### 2025-07-31 Code/Paper_analysisandtables.Rmd
Final cleaning (minus figure folder)

@@
 #correct data types, levels and reference factors
-Data[,c(1:11,16, 19:44, 46:61)] <- lapply(Data[,c(1:11,16, 19:44, 46:61)], factor)
+Data[,c(1:11,16, 19:45, 47:61)] <- lapply(Data[,c(1:11,16, 19:45, 47:61)], factor)
 levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")
@@
 Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
-Data$CONSTITUENTS_OF_EMERGING_CONCERN_RISK_LEVEL_BI <- relevel(Data$CONSTITUENTS_OF_EMERGING_CONCERN_RISK_LEVEL_BI, ref = "NONE")
 Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI <- relevel(Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI, ref = "NONE")
@@
 Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL <- relevel(Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL, ref = "NONE")
-Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI <- relevel(Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI, ref = "NONE")
-Data$EXTREME_WATER_BILL_RISK_LEVEL_BI <- relevel(Data$EXTREME_WATER_BILL_RISK_LEVEL_BI, ref = "NONE")
+Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI <- relevel(Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI, ref = "2")
+Data$EXTREME_WATER_BILL_RISK_LEVEL_BI <- relevel(Data$EXTREME_WATER_BILL_RISK_LEVEL_BI, ref = "2")
 Data$FUNDING_any_failingoratriskonly <- relevel(Data$FUNDING_any_failingoratriskonly, ref = "No")
@@
 
-#MODELS
-
 #Water quality
-WQ1 <- glm(WQ_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+WQ1 <- glm(WQ_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(WQ1)
@@
 ##WQ sub components
-ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(ecoli)
 
-TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(TT)
 
-MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MCL)
@@
 #Affordability 
-Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(Affordability1)
@@
 ## Affordability sub components
-MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MHI)
 
-extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(extreme)
 
-covid <- glm(didnotapplycovid ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+covid <- glm(didnotapplycovid ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(covid)
 
-# TMF
-TMF1 <- glm(TMF_combined_any ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(TMF1)
+# Accessibility
+Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(Accessibility1)
 
-## TMF sub components
-Opcert_violations <- glm(OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(Opcert_violations)
+## Accessibility sub components
+Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(Single_source)
 
-mrviolations <- glm(MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(mrviolations)
+Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(Sourcecapacity_violations)
 
-cash <- glm(DAYS_CASH_ON_HAND_RISK_LEVEL ~ enfranchisement_final + LN_POP + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(cash)
+Service_dis <- glm(Service_disruptions ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
+summary(Service_dis)
 
@@
 
-TMF1OR <- TMF1
-TMF1OR$coefficients <- exp(TMF1OR$coefficients)
-pvaluesTMF1 <- list(summary(TMF1)$coefficients[,4])
+Accessibility1OR <- Accessibility1
+Accessibility1OR$coefficients <- exp(Accessibility1OR$coefficients)
+pvaluesAccessibility1 <- list(summary(Accessibility1)$coefficients[,4])
 
@@
 compositemodellist <- list(WQ1, WQ1OR, Affordability1, Affordability1OR,
-                          TMF1, TMF1OR)
+                          Accessibility1, Accessibility1OR)
 
@@
                                  dep.var.labels = c("Water quality", "Affordability", 
-                                                    "TMF capacity"), 
+                                                    "Accessibility"), 
                                  ci=c(F,T, F, T, F, T), 
@@
                                  p = c(p.valuesWQ1, p.values2, pvaluesAffordability1, p.values2,
-                                       pvaluesTMF1, p.values2),
-                                 omit.stat = c("ll", "aic"),
+                                       pvaluesAccessibility1, p.values2),
+                                 omit.stat = c("ll"),
                                  covariate.labels = c("Limited enfran.", "No enfran.",
-                                                      "Population (log)", "Surface water",
+                                                      "Population", "Surface water",
                                                       "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
@@
                                        pvaluesMCL, p.values2),
-                                 omit.stat = c("ll", "aic"),
+                                 omit.stat = c("ll"),
                                  covariate.labels = c("Limited enfran.", "No enfran.",
-                                                      "Population (log)", "Surface water",
+                                                      "Population", "Surface water",
                                                       "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
@@
                                  p = c(pvaluesMHI, p.values2, pvaluesextreme, p.values2, pvaluescovid, p.values2),
-                                 omit.stat = c("ll", "aic"),
+                                 omit.stat = c("ll"),
                                  covariate.labels = c("Limited enfran.", "No enfran.",
-                                                      "Population (log)", "Surface water",
+                                                      "Population", "Surface water",
                                                       "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
@@
 
-```{r table TMF, results='asis', echo=FALSE}
-#TMF indicators
-Opcert_violationsOR <- Opcert_violations
-Opcert_violationsOR$coefficients <- exp(Opcert_violationsOR$coefficients)
-pvaluesOpcert <- list(summary(Opcert_violations)$coefficients[,4])
+```{r table Accessibility, results='asis', echo=FALSE}
+#Accessibility indicators
+Single_sourceOR <- Single_source
+Single_sourceOR$coefficients <- exp(Single_sourceOR$coefficients)
+pvaluesSinglesource <- list(summary(Single_source)$coefficients[,4])
 
-mrviolationsOR <- mrviolations
-mrviolationsOR$coefficients <- exp(mrviolationsOR$coefficients)
-pvaluesmrviolations <- list(summary(mrviolations)$coefficients[,4])
+Sourcecapacity_violationsOR <- Sourcecapacity_violations
+Sourcecapacity_violationsOR$coefficients <- exp(Sourcecapacity_violationsOR$coefficients)
+pvaluesSourcecapacity_violations <- list(summary(Sourcecapacity_violations)$coefficients[,4])
 
-cashOR <- cash
-cashOR$coefficients <- exp(cashOR$coefficients)
-pvaluescash <- list(summary(cash)$coefficients[,4])
+Service_disOR <- Service_dis
+Service_disOR$coefficients <- exp(Service_disOR$coefficients)
+pvaluesService_dis <- list(summary(Service_dis)$coefficients[,4])
 
-TMFmodellist <- list(Opcert_violations, Opcert_violationsOR, mrviolations, mrviolationsOR, cash, cashOR)
+Accessmodellist <- list(Single_source, Single_sourceOR, Sourcecapacity_violations, Sourcecapacity_violationsOR, Service_dis, Service_disOR)
 
-TMFtable <- capture.output(stargazer(TMFmodellist,
+Accesstable <- capture.output(stargazer(Accessmodellist,
                                  type = 'html', 
@@
                                                    "coefficient (se)", "OR", 
-                                                   "coefficient (se)", "OR", 
                                                    "coefficient (se)", "OR"),
                                  model.numbers = FALSE, 
-                                 dep.var.labels = c("Operator cert. violations", "M&R violations", "Cash on hand"), 
-                                 ci=c(F,T, F, T, F, T, F, T), 
+                                 dep.var.labels = c("Single source", "Source capacity violations", "Service disruptions"), 
+                                 ci=c(F,T, F, T, F, T), 
                                  star.cutoffs = c(0.05, 0.01, 0.001),
@@
                                  notes.append = FALSE,
-                                 p = c(pvaluesOpcert, p.values2, pvaluesmrviolations, p.values2, pvaluescash, p.values2),
-                                 omit.stat = c("ll", "aic"),
+                                 p = c(pvaluesSinglesource, p.values2, pvaluesSourcecapacity_violations, p.values2, pvaluesService_dis, p.values2),
+                                 omit.stat = c("ll"),
                                  covariate.labels = c("Limited enfran.", "No enfran.",
-                                                      "Population (log)", "Surface water",
+                                                      "Population", "Surface water",
                                                       "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
@@
 
-cat(paste(gsub("\\(0.000, 0.000\\)", "", TMFtable), collapse = "\n"), "\n")
+cat(paste(gsub("\\(0.000, 0.000\\)", "", Accesstable), collapse = "\n"), "\n")
 ```
@@
 ```{r descriptives for demofraphics, results='asis', echo=FALSE}
-Data_mini <- Data %>% select(enfranchisement_final, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, TMF_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian)
+
+Data$Source_bi <- NA
+Data$Source_bi <- ifelse(Data$Source == "GW", 1, Data$Source_bi)
+Data$Source_bi <- ifelse(Data$Source == "SW", 0, Data$Source_bi)
+
+Data_mini <- Data %>% select(enfranchisement_final, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian, Source_bi)
 
@@
 TukeyMHI_table <- as.data.frame(TukeyMHI$enfranchisement_final)
-
-Data$Source_bi <- NA
-Data$Source_bi <- ifelse(Data$Source == "GW", 1, Data$Source_bi)
-Data$Source_bi <- ifelse(Data$Source == "SW", 0, Data$Source_bi)
 

### 2025-07-31 Code/sensitivity_tests_march2025_tables_profit.Rmd
Final cleaning (minus figure folder)

@@
 #LOAD AND READY DATA
-Data <- read.csv(here::here("Data_processed/Systemperformancedata_RR.csv"))
+Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
 Data <- Data[,-1]

### 2025-07-31 Code/Models.R

@@ -8,7 +8,7 @@ library(vtable)
 library(sjPlot)
 
 #LOAD AND READY DATA
-Data <- read.csv(here::here("Data_processed/Systemperformancedata_RR.csv"))
+Data <- read.csv(here::here("Data_processed/Systemperformancedata.csv"))
 Data <- Data[,-1]
 Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
 Data$Final_inst_update <- as.factor(Data$Final_inst_update)

### 2025-07-31 Code/sensitivity_tests_march2025.R

@@ -4,7 +4,9 @@
 #subbing enfranchisement for public private 
 Data$public <- relevel(Data$public, ref = "Yes")
 
-#Water quality
+#Water quality\
+library(car)
+library(regclass)
 summary(WQ1)
 
 WQ1P <- glm(WQ_combined_any ~  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)

### 2025-07-31 Code/Figures_PP.R
Final cleaning of figures

@@ -30,7 +30,7 @@ plotCombined <- Probabilities %>%
   facet_wrap(~Key, scales = "free", nrow = 1, ncol = 3) +
   cowplot::theme_half_open(); plotCombined
 
-ggsave("Fig2_RR.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
+ggsave("Fig2.jpeg", plotCombined, path = "Figures/", height = 6, width = 7.487, units = "in", dpi = 720)
 
 #water quality subindicators figure
 Effect_WQ_ecoli <- predictions(ecoli,
@@ -122,7 +122,7 @@ plotCombined_MASSIVE <- Probabilities_MASSIVE %>%
   facet_wrap(~Category+Key, scales = "free", nrow = 4, ncol = 3) +
   cowplot::theme_half_open(); plotCombined_MASSIVE
 
-ggsave("Fig3_RR.jpeg", plotCombined_MASSIVE, path = "Figures/", height = 6, width = 9, units = "in", dpi = 720)
+ggsave("Fig3.jpeg", plotCombined_MASSIVE, path = "Figures/", height = 6, width = 9, units = "in", dpi = 720)
 
 #Make bar graph for enfranchisement variables
 

### 2025-09-01 Code/sensitivity_tests_march2025_tables_specialization.Rmd
Final code clean up for paper publication

@@
 ```{r setup, include=FALSE, echo=FALSE}
-#MARKDOWN SETTINGS
 #knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Documents/Projects/R_Projects/Votingrights_systemperformance", warning=FALSE, message=FALSE, include=FALSE)

### 2025-09-01 Code/Data_compilation_systemperformance.R

@@ -7,7 +7,7 @@ library(plyr)
 # Create sample and governance variables ----------------------------------------------------
 
 ###Load data and initial manipulations
-CWSgov <- read.csv(here::here("Data_raw/Dobbin_Fencl_McBride_2023_CA_CWS_institutional_type_update.csv"), header=T, na.strings=c("","NA"))
+CWSgov <- read.csv(here::here("Data_raw/Dobbin_Fencl_McBride_2023_CA_CWS_institutional_type_update.csv"), header=T, na.strings=c("","NA")) #Dobbin, Fencl and McBride 2023 California water system institutional update dataset from Dryad
 CWSgov$Primacy_FINAL <- as.factor(CWSgov$Primacy_FINAL)
 CWSgov <- CWSgov %>% filter(Primacy_FINAL != "EPA Region 9") #get rid of EPA regulated Tribal water systems
 CWSgov <- CWSgov %>% filter(Final_inst_update != "Tribal Government") #Get rid of three remaining systems that are Tribal but state regulated
@@ -18,7 +18,7 @@ CWSgov$Inst_Subtype  <- as.factor(CWSgov$Inst_Subtype)
 
 ### create new variables for ancillary/non ancillary, public/private and residential enfranchisement
 
-## To focus on residential systems where residents are served and could have a role in governance/decision-making, identify which systems are "residential" this is mostly synonymous to non-ancillary except with this variable maintaining Mobile Home Parks which are generally considered to be ancillary and yet in other ways these systems are more similar to non-ancillary systems in that they serve a stable residential base. Moreover, California has seen significant organizing among Mobile Home Park communities in recent years around drinking water issues. 
+## To focus on residential systems where residents are served and could have a role in governance/decision-making, identify which systems are "residential" this is mostly synonymous to non-ancillary except with this variable maintaining Mobile Home Parks which are generally considered to be ancillary and yet in other ways these systems are more similar to non-ancillary systems in that they serve a stable residential base. Moreover, California has seen significant organizing among Mobile Home Park communities in recent years around drinking water issues and they are one of the most common water providers in the state. 
 CWSgov$Residential <- ifelse(CWSgov$Final_inst_update == "Community Services District" |
                                CWSgov$Final_inst_update == "County Service Area" | 
[…]

### 2025-09-01 Code/Figures_Map.R

@@ -1,5 +1,5 @@
 #Map systems by enfranchisement type.
-## NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Data_compilation_SES.R script
+##NOTE THIS SCRIPT REQUIRES FIRST RUNNING THE Data_compilation_SES.R script
 
 #Load libraries
 library(tidyverse)
@@ -63,24 +63,6 @@ Demo <- ggplot() +
 
 ggsave('Figures/Fig1b.png', width = 4, height = 4, dpi = 720, bg='#ffffff')
 
-#Make a county map for a more simple smaller/county
-
-caCountiesTmp_Yolo <- caCountiesTmp %>% filter(NAME == "Yolo")
-Yolo <- Data_geo %>% filter(Data_geo$COUNTY_FIXED == "YOLO")
-
-Demo5 <- ggplot() +
-  geom_sf(data = caCountiesTmp_Yolo) +
-  geom_sf(mapping = aes(colour = enfranchisement_final, geometry = geometry), data = Yolo, 
-          inherit.aes = FALSE) +
-  theme_void() +
-  theme(legend.position = "bottom", 
-        legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(10,10,10,10)) +
-  labs(color = "Enfranchisement type") +
-  scale_color_brewer(palette = "Dark2") +
-  theme(panel.background=element_blank()); Demo5
-
-ggsave('Figures/Demo_map4.png', width = 6, height = 3, dpi = 300, bg='#ffffff')
 
 #I think Mendocino is best, Combine into panel figure
-Combined <- grid.arrange(Enfranchisement_Type, Demo4, nrow=1)
+Combined <- grid.arrange(Enfranchisement_Type, Demo, nrow=1)

### 2025-09-01 Code/Models.R

@@ -1,6 +1,6 @@
 # Models
-## Script includes creation of the "Fulldataset.csv" file which is the mater data for this project
-## This code should match Paper_analysisandtables.Rmd which is used to generate tables
+## Script includes creation of the "Fulldataset.csv" file which is the master data for this project
+## This code should match Paper_analysisandtables.Rmd which is used to generate pub ready tables
 
 #LOAD LIBRARIES
 library(tidyverse)
@@ -53,13 +53,6 @@ Demographics <- Demographics[,-c(1,3)]
 Demographics$PWSID <- as.factor(Demographics$PWSID)
 
 Data <- left_join(Data, Demographics, by = "PWSID")
-
-#Remove large systems since they don't have most DV variables anyways. NOT DOING THIS NOW. REMOVES MORE DATA THAN NEEDED
-#Data <- Data %>% filter(POPULATION > 24 | SERVICE_CONNECTIONS > 14) #To remove wholesalers
-#Data <- Data %>% filter(SERVICE_CONNECTIONS <3301) #To remove large systems
-#Data <- Data %>% filter(POPULATION <100000) #To get rid of another wholesaler
-#Data <- Data %>% filter(POPULATION != 0) #To get rid of more wholsalers
-#reduced overall dataset from 2405 to 1953
 
 write.csv(Data, "Data_processed/Fulldataset_RR.csv")
 

### 2025-09-01 Code/sensitivity_tests_march2025.R

@@ -1,10 +1,10 @@
 #Sensitivity tests conducted March 2025
 ## MUST RUN models script before this one
 
-#subbing enfranchisement for public private 
+#################subbing enfranchisement for public private 
 Data$public <- relevel(Data$public, ref = "Yes")
 
-#Water quality\
+#Water quality
 library(car)
 library(regclass)
 summary(WQ1)
@@ -15,7 +15,7 @@ VIF(WQ1P)
 
 WQ1P2 <- glm(WQ_combined_any ~ enfranchisement_final +  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(WQ1P2)
-VIF(WQ1P2)
+VIF(WQ1P2) #VIF way too high
 
 ##WQ sub components
 
@@ -24,36 +24,26 @@ VIF(ecoli)
 
 ecoliP <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(ecoliP)
-
-ecoliP2 <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final +  public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(ecoliP2)
-VIF(ecoliP2)
+VIF(ecoliP)
 
 summary(TT)
 
 TTP <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(TTP)
-
-TTP2 <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
-summary(TTP2)
-
+VIF(TTP)
 
 summary(MCL)
 
 MCLP <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ public + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
 summary(MCLP)
-
[…]
